### 2024-12-21 text-paper/paper.tex
first commit

\documentclass[10pt]{article}
\usepackage{multicol}
\usepackage{authblk}
\usepackage{mathptmx}
\usepackage{graphicx}
\setlength{\columnsep}{5mm}
\usepackage{blindtext}
\usepackage{geometry}
\usepackage{url}

\geometry{
  a4paper,
  left = 20mm,
  top = 2mm,
  right = 20mm
}

\title{\textbf{Parallel dragonfly algorithm} %
  \\[2ex] \large High Performance Computing for Data Science course 2024/2025}

\author[1]{Mattia Santaniello}
\author[2]{Alessio Zeni}
\affil{University of Trento}

\begin{document}

\maketitle

\begin{multicols}{2}[\fontsize{9}{9}\section*{Abstract} \textbf{Dragonfly algorithm is a new kind of approach used to solve optimization problems through the emulation of dragonflies movements, but it lacks scalability. In this paper the authors try to set up an approach based on parallelization thanks to the use of API for high performance computing like MPI and OpenMP, then there will be a comparison of execution times between presented implementation and the classic approach based on serial programming, in order to see if parallelization could lead to a performance improvement and possible future extensions.}\newline]

\section{Introduction}

Recent years have seen the raise and evolution of swarm intelligent algorithms, based on meta-heuristic approach where single individuals of a complex population follow an exploration-exploitation methodology to find the best solution in their local search space; each member of the group work independently from the others, leading to a subdivision of the labour which decrease the overall amount of execution time, working as decentralized and parallel units; but also, due to the fact that they mimic the behaviour of biological life forms, it is possible then that they could generate random patterns through the interaction with each other.  The dragonfly algorithm has been proposed for the first time in 2016, and its purpose is to simulate the behaviour of a dragonfly population according to the principles of swarm intelligent algorihms seen previously. It has gained more and more popularity: it is in estimated that during 2019 more than 300 scientific papers cited the algorithm \cite{DASurvey}, due to the fact that it has demonstrated some characteristics similar to other algoritms of the same kind, however in contrast with other swarm intelligent algorithms however it present less probability to be trapped into a local optima \cite{DAReview}. This paper aims to define a new approach based on popular parallel APIs, and also trying to understand if parallelization could lead to a significant speed up in performances that can be considered by future researches on this algorithm

\subsubsection*{How the algorithm works}
The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$; then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program. This procedure is simulated through the observation of repetitive patterns observed in the behaviour of dragonflies, which are:

\begin{itemize}
  \item Alignment of their velocity with other members of the same group, given by the formula  $A_i = \frac{\sum^M_{j=1}V_j}{M}$, where $V_j$ is the velocity of the j-th neighbour

  \item Separation of single members in order to avoid collision, and it is retrieved by the equation  $S_i = - \sum^M_{j=1}{P - P_j}$, where $P_j$ is the position of the j-th neighbour and $P$ the current individual position.

  \item Cohesion towards the center of the group, given by $C_i = \frac{\sum^M_{j=1}P_j}{M} - P$ 
  \item Attraction to a common source food, expressed by the formula $ F_i = F^+ - P$, where $F^+$ is the position of the current prey and $P$ is the position of the current dragonfly
  \item Distraction of enemies, represented by the equation $ E_i = E^- - P$ where $E^-$ is the position of the current enemy 
\end{itemize}

\noindent Once all the parameters have been calculated we can now update the current step vector of each dragonfly using the following equation: $$\Delta P^{t+1}_{i} = (sS_i + aA_i + fF_i + cC_i) + \omega\Delta P^t_i$$ where $s,a,f,c$ are respectevly the separation weight, the alignment weight, the attraction weight, the cohesion weight and $\omega$ is the inertia weight, all of them are randomly chosen in the interval $[0,1]$. When all step vectors have been updated we can now set the new position of the dragonflies using this equation, which takes the new step vector as an input: $$P^{t+1}_{i} = P_i^t + \Delta P_{i}^{t+1}$$. However if there is a dragonfly which does not have any neighbour then we assume that the overall population moves in a random way: the most common method to do that is to use a random walk function. The previous formula then must be modified into this one here: $$P^{t+1}_i = P_i^t + Levy(d)$$ where $d$ is the dimension of the position vector, and $Levy(d)$ is the Levy flight function, which is calculated in this way: $$Levy(d) = 0.01\times\frac{r_1 \times \sigma}{|r_2|^{\frac{1}{\beta}}}$$ where $r1$ and $r2$ are two random numbers uniformally distributed in the interval $[0,1]$, and $\sigma$ is a vector calculated using the following formula: $$\sigma = \frac{sin(\frac{\beta\pi}{2})\times\Gamma(1+\beta)}{\Gamma(\frac{\beta+1}{2})\times \beta \times 2^{(\beta-1)/2}}$$ where $\Gamma(x)$ is the gamma function \cite{WikiGamma}, and $\beta$ is a constant.Another valid approach is to apply, instead of Levy flight, the brownian motion model, used especially in gas and particle analysis: this seems to return better performances \cite{BDragonfly}.

\section{first approach}

\end{multicols}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}

### 2024-12-21 code/build.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+make && gdb ./dragonfly-serial

### 2024-12-21 code/main.c

@@ -1,0 +1,176 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <stdbool.h>
+#include <math.h>
+
+#define POP_SIZE 1000
+#define RAND_FLOAT(N) ((float)rand()/(float)RAND_MAX)*N
+#define ZEROS(N) init_array(N,0);
+
+// take timing not including IO
+
+typedef struct{
+  float x;
+  float y;
+} Coord;
+
+
+Coord* init_array(int dim,int range_max){
+	Coord *arr = (Coord*)malloc(dim*sizeof(Coord));
+
+	for(int i=0; i<POP_SIZE; i++){
+      arr[i] = (Coord){.x = RAND_FLOAT(range_max), .y = RAND_FLOAT(range_max)}; 
+	}
+
+	return arr;
+}
+
+
+float distance(Coord p1,Coord p2){
+  double dx = p1.x - p2.x;
+  double dy = p1.y - p2.y;
+
+  return sqrt(pow(dy,2)+pow(dx,2)); 
+}
+
+
+bool get_updates(Coord* position,Coord* step,Coord* food,Coord* enemy_pos,float* weights,Coord* updates){
+  bool neighbours = false;
+  Coord S;
+  Coord A;
+  Coord C;
+  Coord F;
+  Coord E;
+  int nr_neighbours;
+
+
+  float eps = 10e-2;
+
+  for(int dragonfly=0; dragonfly<POP_SIZE; dragonfly++){
+    S = (Coord){.x = 0,.y=0};
+    A = (Coord){.x = 0,.y=0};
+    C = (Coord){.x = 0,.y=0};
+    F = (Coord){.x = 0,.y=0};
+    E = (Coord){.x = 0,.y=0};
+    nr_neighbours = 0;
+
+    for(int k=0; k<POP_SIZE; k++){
+
+      if(k != dragonfly && distance(position[dragonfly],position[k]) < eps){
+            S.x += position[k].x - position[dragonfly].x;
+            S.y += position[k].y - position[dragonfly].y;
+
+            A.x += step[k].x;
+            A.y += step[k].y;
+
+            C.x += position[k].x;
+            C.y += position[k].y;
+
+            nr_neighbours++;
+        }
+
+    }
+
+    F.x = food[dragonfly].x - position[dragonfly].x;
+    F.y = food[dragonfly].y - position[dragonfly].y;
+
[…]

### 2025-02-10 code/utils-parallel.c
fixed memory leak

@@ -1,10 +1,13 @@
 #include "dragonfly-common.h"
 #include "utils-special.h"
 #include <omp.h>
-#include <stdio.h>
 
-#define NR_THREADS 4
+//#define NR_THREADS 4
+unsigned int NR_THREADS =4;
 
+void set_thread_number(unsigned int N){
+	  NR_THREADS = N;
+}
 void dragonfly_compute_step(Dragonfly *d, float *average_speed,
                             float *cumulated_pos, float *food, float *enemy,
                             unsigned int N){

### 2025-02-10 code/utils-special.h

@@ -10,4 +10,6 @@ void dragonfly_compute_step(Dragonfly *d, float *average_speed,
                             unsigned int N);
 void computation_accumulate(ComputationStatus *message, Dragonfly *d, float* best, float* best_fitness);
 
+void set_thread_number(unsigned int N);
+
 #endif

### 2025-02-10 code/utils.c

@@ -1,5 +1,4 @@
 #include <math.h>
-#include <stdio.h>
 #include <stdlib.h>
 
 #include "dragonfly-common.h"

### 2025-02-11 code/bin_training_mpich.c
adding runtime threads

@@ -42,7 +42,7 @@ float eval(float *wi, unsigned int *seed, unsigned int d) {
       .wl = {wi[10], wi[11]},
       .ll = {wi[12], wi[13]},
   };
-  Parameters p = {.n = 128, .dim = 10, .chunks = 64, .iterations = 80};
+  Parameters p = {.n = 128, .dim = 10, .chunks = 64, .iterations = 80, .threads_per_process = 1};
 
   Fitness fitness = shifted_fitness;
   float avg = 0.0;
@@ -87,6 +87,7 @@ int main(int argc, char **argv) {
   srand(time(NULL) + rank);
   Fitness fitness = eval;
   Parameters p = parameter_parse(argc, argv);
+  printf("%d\n", p.threads_per_process);
 
   float wi[14] = {
       0.000000,

### 2025-02-11 code/dragonfly-common.c

@@ -1,7 +1,6 @@
 #include "dragonfly-common.h"
 #include "stdlib.h"
 #include "string.h"
-#include "utils-special.h"
 #include "utils.h"
 
 #include <stdbool.h>
@@ -14,7 +13,7 @@
 // n, chunks, iterations, dim
 Parameters parameter_parse(int argc, char* argv[]){
   Parameters p;
-  if(argc!=5){
+  if(argc!=6){
     fprintf(stderr, "invalid parameter count: expected n, chunks, iterations, dimensions");
     exit(-1);
   }
@@ -22,7 +21,10 @@ Parameters parameter_parse(int argc, char* argv[]){
   p.chunks=atoi(argv[2]);
   p.iterations=atoi(argv[3]);
   p.dim=atoi(argv[4]);
-  if(p.n==0 || p.chunks==0 || p.iterations==0 || p.dim==0){
+
+  p.threads_per_process=atoi(argv[5]);
+
+  if(p.n==0 || p.chunks==0 || p.iterations==0 || p.dim==0 || p.threads_per_process==0){
     fprintf(stderr, "invalid parameter they must be all bigger than 1 (and integers)");
     exit(-1);
   }
@@ -218,6 +220,9 @@ float *dragonfly_compute(Parameters p, Weights w, Fitness fitness, unsigned int 
       exit(-1);
     }
   #endif
+  if (p.threads_per_process>1){
+
+  }
   // compute start and end rank
   unsigned int start_chunk = rank_id*(p.chunks/threads) + min(rank_id, p.chunks%threads);
   unsigned int current_chunks = p.chunks/threads + (rank_id<(p.chunks%threads));
@@ -264,7 +269,7 @@ float *dragonfly_compute(Parameters p, Weights w, Fitness fitness, unsigned int 
 
     // compute avarage speed and positions
     for(unsigned int j=0; j<current_chunks; j++){
-      computation_accumulate(&message.status[j+start_chunk], &d[j], best, &best_fitness);
+      computation_accumulate(&message.status[j+start_chunk], &d[j], best, &best_fitness, p.threads_per_process);
     }
     message.start_chunk=start_chunk;
     message.end_chunk=end_chunk;
@@ -286,7 +291,7 @@ float *dragonfly_compute(Parameters p, Weights w, Fitness fitness, unsigned int 
         if(start_chunk<=j+k && j+k<end_chunk){
           dragonfly_compute_step(&d[j+k-start_chunk], message.status[j].cumulated_speeds,
[…]

### 2025-02-11 code/dragonfly-common.h

@@ -63,7 +63,7 @@ typedef struct{
 
 
 typedef struct {
-  unsigned int n, chunks, iterations, dim;
+  unsigned int n, chunks, iterations, dim, threads_per_process;
 } Parameters;
 
 Parameters parameter_parse(int argc, char *argv[]);

### 2025-02-11 code/utils-parallel.c

@@ -1,172 +1,0 @@
-#include "dragonfly-common.h"
-#include "utils-special.h"
-#include <omp.h>
-
-//#define NR_THREADS 4
-unsigned int NR_THREADS =4;
-
-void set_thread_number(unsigned int N){
-	  NR_THREADS = N;
-}
-void dragonfly_compute_step(Dragonfly *d, float *average_speed,
-                            float *cumulated_pos, float *food, float *enemy,
-                            unsigned int N){
-unsigned int rest = d->N % NR_THREADS;
-unsigned int ratio = d->N / NR_THREADS;
-unsigned int dimensions = d->dim;
-
-#pragma omp parallel num_threads(NR_THREADS)
-{
-
-	
-	unsigned int rank = omp_get_thread_num(); 
-	unsigned int base = ratio*rank;
-	unsigned int limit = ratio*(rank+1);
-
-
-	float S;
-	float A;
-	float C;
-	float F;
-	float E;
-	float levy;
-
-	float* cur_pos;
-	float* cur_speed;
-
-    for(unsigned int j=base; j<limit;j++){
-	  cur_pos = d->positions+j*dimensions;
-	  cur_speed = d->speeds+j*dimensions;
-
-
-      //compute speed = sSi + aAi + cCi + fFi + eEi + w
-
-	  for(unsigned int i = 0; i<dimensions; i++){
-		  S = (cumulated_pos[i]/((float)N)) - cur_pos[i];
-		  A = average_speed[i];
-		  C = (cumulated_pos[i]/(float)N)-cur_pos[i];
-		  F = food[i]-cur_pos[i];
-		  E = enemy[i]+cur_pos[i];
-		  levy = RAND_FLOAT(1.0,&d->seed); 
-		  
-
-		  cur_speed[i] *= d->w.w;
-		  cur_speed[i] += d->w.s*S;
-		  cur_speed[i] += d->w.a*A;
-		  cur_speed[i] += d->w.c*C;
-		  cur_speed[i] += d->w.f*F;
-		  cur_speed[i] += d->w.e*E;
-		  cur_speed[i] += levy;
-
-		  cur_pos[i] += cur_speed[i];
-	  }
-
-    }
-}
-	if(rest != 0){
-		unsigned int r_base = ratio*NR_THREADS;
-		unsigned int r_end = d->N-1;
-
-		for(unsigned int j = r_base; j<=r_end; j++){
[…]

### 2025-02-11 code/utils-serial.c

@@ -1,87 +1,0 @@
-#include "dragonfly-common.h"
-#include"utils-special.h"
-
-void dragonfly_compute_step(Dragonfly *d, float *average_speed,
-                            float *cumulated_pos, float * food, float * enemy, unsigned int N) {
-  unsigned int dim = d->dim;
-  // for each dragonfly
-  for (unsigned int j = 0; j < d->N; j++) {
-    float *cur_pos = d->positions + dim * j;
-    float *cur_speed = d->speeds + dim * j;
-
-    // compute separation: Si = -sumall(X-Xi)
-    memcpy(d->S, cumulated_pos, sizeof(float) * dim);
-    scalar_prod_assign(d->S, 1.0/(float)N, dim);
-    sub_assign(d->S, cur_pos, dim);
-    scalar_prod_assign(d->S, d->w.s, dim);
-
-    // compute alignament: Ai = avarage(Vi)
-    memcpy(d->A, average_speed, sizeof(float) * dim);
-    scalar_prod_assign(d->A, d->w.a, dim);
-
-    // compute cohesion: Ci = avarage(Xi)-X
-    memcpy(d->C, cumulated_pos, sizeof(float) * dim);
-    scalar_prod_assign(d->C, 1.0 / (float)N, dim);
-    sub_assign(d->C, cur_pos, dim);
-    scalar_prod_assign(d->C, d->w.c, dim);
-
-    // food attraction: Fi=X_food - X
-    memcpy(d->F, food, sizeof(float) * dim);
-    sub_assign(d->F, cur_pos, dim);
-    scalar_prod_assign(d->F, d->w.f, dim);
-
-    // enemy repulsion: E=X_enemy+X
-    memcpy(d->E, enemy, sizeof(float) * dim);
-    sum_assign(d->E, cur_pos, dim);
-    scalar_prod_assign(d->E, d->w.e, dim);
-
-    brownian_motion(d->levy, dim, &d->seed);
-    
-
-    // compute speed = sSi + aAi + cCi + fFi + eEi + w
-    scalar_prod_assign(cur_speed, d->w.w, dim);
-    sum_assign(cur_speed, d->E, dim);
-    sum_assign(cur_speed, d->F, dim);
-    sum_assign(cur_speed, d->C, dim);
-    sum_assign(cur_speed, d->A, dim);
-    sum_assign(cur_speed, d->S, dim);
-    sum_assign(cur_speed, d->levy, dim);
-    
[…]

### 2025-02-11 code/utils-special.h

@@ -1,15 +1,0 @@
-#ifndef DA_PARALLEL
-#define DA_PARALLEL
-#include"dragonfly-common.h"
-#include <string.h>
-#include <stdlib.h>
-#include"utils.h"
-
-void dragonfly_compute_step(Dragonfly *d, float *average_speed,
-                            float *cumulated_pos, float *food, float *enemy,
-                            unsigned int N);
-void computation_accumulate(ComputationStatus *message, Dragonfly *d, float* best, float* best_fitness);
-
-void set_thread_number(unsigned int N);
-
-#endif

### 2025-02-11 code/utils.c

@@ -22,7 +22,8 @@ void* my_malloc(size_t size, const char *file, int line, const char *func)
 }*/
 #include "utils.h"
 
-float *init_array(unsigned int dimensions, float range_max, unsigned int *seed) {
+float *init_array(unsigned int dimensions, float range_max,
+                  unsigned int *seed) {
   float *arr = malloc(dimensions * sizeof(float));
   for (unsigned int i = 0; i < dimensions; i++) {
     arr[i] = RAND_FLOAT(range_max, seed);
@@ -54,8 +55,8 @@ void scalar_prod_assign(float *dest, float val, unsigned int size) {
   }
 }
 
-float dot_product(float * a, float * b, unsigned int size){
-  float ret =0.0;
+float dot_product(float *a, float *b, unsigned int size) {
+  float ret = 0.0;
   for (unsigned int i = 0; i < size; i++) {
     ret += a[i] * b[i];
   }
@@ -64,37 +65,39 @@ float dot_product(float * a, float * b, unsigned int size){
 
 float length(float *inp, unsigned int size) {
   float ret = dot_product(inp, inp, size);
-  
+
   return sqrt(ret);
 }
 
-//Gram-Schmidt orthonormal matrix
-void init_matrix(float *inp, float range_max, unsigned int dim, unsigned int *seed){
-  float * tmp = malloc(sizeof(float)*dim);
-  for(unsigned int row=0; row<dim; row++){
+// Gram-Schmidt orthonormal matrix
+void init_matrix(float *inp, float range_max, unsigned int dim,
+                 unsigned int *seed) {
+  float *tmp = malloc(sizeof(float) * dim);
+  for (unsigned int row = 0; row < dim; row++) {
     for (unsigned int col = 0; col < dim; col++) {
-      inp[row*dim+col] = RAND_FLOAT(range_max, seed);
+      inp[row * dim + col] = RAND_FLOAT(range_max, seed);
     }
 
-    //make it perpendicular
-    for(unsigned int prow=0; prow<row;prow++){
-      float cur=dot_product(inp+ row*dim, inp+prow*dim, dim);
-      cur/=dot_product(inp+ row*dim, inp+prow*dim, dim);
-      memcpy(tmp, inp+prow*dim, dim*sizeof(float));
+    // make it perpendicular
+    for (unsigned int prow = 0; prow < row; prow++) {
[…]

### 2025-02-11 code/utils.h

@@ -25,4 +25,10 @@ float shifted_fitness(float *inp, unsigned int* seed, unsigned int dim);
 void init_shifted_fitness(float *tmp, float * rotation, float * shift, Fitness fitness);
 void init_matrix(float *inp, float range_max, unsigned int dim, unsigned int *seed);
 
+
+void dragonfly_compute_step(Dragonfly *d, float *average_speed,
+    float *cumulated_pos, float *food, float *enemy,
+    unsigned int N, unsigned int NUM_THREADS);
+void computation_accumulate(ComputationStatus *message, Dragonfly *d, float* best, float* best_fitness, unsigned int NUM_THREADS);
+
 #endif

### 2025-02-11 code/bin_training_mpich.c
adding runtime threads

@@ -1,6 +1,7 @@
 #include "dragonfly-common.h"
 #include "utils.h"
 #include <mpi.h>
+#include <omp.h>
 #include <stdbool.h>
 #include <stdio.h>
 #include <stdlib.h>
@@ -87,7 +88,7 @@ int main(int argc, char **argv) {
   srand(time(NULL) + rank);
   Fitness fitness = eval;
   Parameters p = parameter_parse(argc, argv);
-  printf("%d\n", p.threads_per_process);
+  omp_set_num_threads(p.threads_per_process);
 
   float wi[14] = {
       0.000000,

### 2025-02-11 code/utils.c

@@ -166,93 +166,98 @@ float shifted_fitness(float *inp, unsigned int *seed, unsigned int dim) {
 
 #ifdef USE_OPENMP
 #include <omp.h>
-void dragonfly_compute_step_multithread(Dragonfly *d, float *average_speed,
-                                        float *cumulated_pos, float *food,
-                                        float *enemy, unsigned int N,
-                                        unsigned int NR_THREADS) {
+
+#endif
+
+void inner_dragonfly_step(Dragonfly *d, float *average_speed,
+                          float *cumulated_pos, float *food, float *enemy,
+                          unsigned int N, unsigned int base, unsigned int limit,
+                          unsigned int random_seed) {
+  unsigned int dimensions = d->dim;
+  float S;
+  float A;
+  float C;
+  float F;
+  float E;
+  float levy;
+
+  float *cur_pos;
+  float *cur_speed;
+
+  for (unsigned int j = base; j < limit; j++) {
+    unsigned random = random_seed + j;
+    cur_pos = d->positions + j * dimensions;
+    cur_speed = d->speeds + j * dimensions;
+
+    // compute speed = sSi + aAi + cCi + fFi + eEi + w
+
+    for (unsigned int i = 0; i < dimensions; i++) {
+      S = (cumulated_pos[i] / ((float)N)) - cur_pos[i];
+      A = average_speed[i];
+      C = (cumulated_pos[i] / (float)N) - cur_pos[i];
+      F = food[i] - cur_pos[i];
+      E = enemy[i] + cur_pos[i];
+      levy = RAND_FLOAT(1.0, &random);
+
+      cur_speed[i] *= d->w.w;
+      cur_speed[i] += d->w.s * S;
+      cur_speed[i] += d->w.a * A;
+      cur_speed[i] += d->w.c * C;
+      cur_speed[i] += d->w.f * F;
+      cur_speed[i] += d->w.e * E;
+      cur_speed[i] += levy;
+
+      cur_pos[i] += cur_speed[i];
+    }
+  }
+}
+void dragonfly_compute_step(Dragonfly *d, float *average_speed,
+                            float *cumulated_pos, float *food, float *enemy,
+                            unsigned int N, unsigned int NR_THREADS) {
+  unsigned int base_random = rand_r(&d->seed);
+
+  // if more than 1 thread
   unsigned int rest = d->N % NR_THREADS;
   unsigned int ratio = d->N / NR_THREADS;
[…]

### 2025-02-11 code/dragonfly-common.h
too small

@@ -1,7 +1,7 @@
 #ifndef DA_COMMON
 #define DA_COMMON
 #define MESSAGE_SIZE 16
-#define MAX_CHUNKS 1024
+#define MAX_CHUNKS 4048
 
 typedef struct {
   float sl[2], al[2], cl[2], fl[2], el[2], wl[2], ll[2];

### 2025-02-11 text-paper/paper.tex
updated report

@@
 \usepackage{url}
+\usepackage{listings}
+\usepackage{xcolor}
+\usepackage{float}
+\usepackage{longtable}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\small,
+    columns=flexible,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=2pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+
+\lstset{style=mystyle}
 
@@
 
-\subsubsection*{How the algorithm works}
+\subsection*{How the algorithm works}
 The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$; then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program.
@@
 
+Now that we have seen the overall procedure it is possible to write a naive implementation of the algorithm in order to 
+have a base for further developments. An example of the implementation can be observed in the following code snippet:
+
+\begin{lstlisting}[language=C,caption={first implementation of the dragonfly algorithm}]
+  void dragonfly_algorithm(Dragonfly *d,
+                          float *average_speed, 
+                          float *cumulated_pos, 
+                          float * food, 
+                          float * enemy, 
+                          unsigned int N) {
+
+    unsigned int dim = d->dim;
+      // for each dragonfly
+    for (unsigned int j = 0; j < d->N; j++) {
+      float *cur_pos = d->positions + dim * j;
+      float *cur_speed = d->speeds + dim * j;
+
+      // compute separation: Si = -sumall(X-Xi)
+      memcpy(d->S, cumulated_pos, sizeof(float) * dim);
+      scalar_prod_assign(d->S, 1.0/(float)N, dim);
+      sub_assign(d->S, cur_pos, dim);
+      scalar_prod_assign(d->S, d->w.s, dim);
+
+      // compute alignament: Ai = avarage(Vi)
+      memcpy(d->A, average_speed, sizeof(float) * dim);
+      scalar_prod_assign(d->A, d->w.a, dim);
+
+      // compute cohesion: Ci = avarage(Xi)-X
+      memcpy(d->C, cumulated_pos, sizeof(float) * dim);
+      scalar_prod_assign(d->C, 1.0 / (float)N, dim);
+      sub_assign(d->C, cur_pos, dim);
+      scalar_prod_assign(d->C, d->w.c, dim);
+
+      // food attraction: Fi=X_food - X
+      memcpy(d->F, food, sizeof(float) * dim);
+      sub_assign(d->F, cur_pos, dim);
+      scalar_prod_assign(d->F, d->w.f, dim);
+
+      // enemy repulsion: E=X_enemy+X
+      memcpy(d->E, enemy, sizeof(float) * dim);
+      sum_assign(d->E, cur_pos, dim);
+      scalar_prod_assign(d->E, d->w.e, dim);
+
+      brownian_motion(d->levy, dim, &d->seed);
+
+      // compute speed = sSi + aAi + cCi + fFi + eEi + w
+      scalar_prod_assign(cur_speed, d->w.w, dim);
+      sum_assign(cur_speed, d->E, dim);
+      sum_assign(cur_speed, d->F, dim);
+      sum_assign(cur_speed, d->C, dim);
+      sum_assign(cur_speed, d->A, dim);
+      sum_assign(cur_speed, d->S, dim);
+      sum_assign(cur_speed, d->levy, dim);
+
+      // update current pos
+      sum_assign(cur_pos, cur_speed, dim);
+    }
+
+    // update weights
+    weights_step(&d->w);
+  }
+\end{lstlisting}
+
+\noindent After the initialization of velocities and positions of agents,
+which will be represented by two separated matricies with the same dimension equal to the number of members of our
+population times the dimension of our problem, the algorithm starts to execute a for loop with a fixed amount of iteration stored in the variable $d\rightarrow N$, 
+and will repeatetly compute the following steps:
+\begin{itemize}
+  \item retrieve current speed and position of the j-th dragonfly
+  \item compute separation, alignment, cohesion, food attraction and enemy distraction vectors,
+  stored into arrays placed in the struct Dragonfly
+  \item calculate product between weigths and arrays
+  \item update speed and position the dragonfly considering also levy flight
+\end{itemize}
+\subsection*{chunks}
+We can improve this approach introducing groupments of dragonflies into chunks, 
+allowing us to calculate local minima of different areas and compare them in order
+to get as close as possible to the global minima. which will be considered
+autonomous agents that will calculate local minima of our fitness function, updating then the 
+global status of by sending messages to all other dragonfly chunks.
+
+\begin{figure}[H]
+  \includegraphics[scale=0.3]{chunks.png}
+  \caption{illustration of communication between chunks}
+\end{figure}
+
+\section{Performace tests}
+
+The serial implementation seen above can be easily tried out using some test function 
+with known global minima in order to check the overall performances. We need to 
+report time execution for each test function, using the input dimension as independent variable;
+since however we have multiple input parameters to compare, we also have to decide which one will 
+be considered during the performance evaluation.
+
+\begin{table}[H]
+\begin{tabular}{| c | c | c | c |}
+  \hline
+  \textit{Nr of iterations} & \textbf{sphere [ms]} & \textbf{rosenblock [ms]} & \textbf{rastrigin [ms]}\\
+  \hline
+  10 \\
+  \hline
+
+  100 \\
+  \hline
+
+  1000 \\
+  \hline
+
+  10000 \\ 
+  \hline
+
+  100000 \\
+  \hline
+\end{tabular}
+
+\end{table}
+
+\begin{table}[H]
+  \begin{tabular}{| c | c | c | c |}
+    \hline
+    \textit{Nr of iterations} & \textbf{sphere [ms]} & \textbf{rosenblock [ms]} & \textbf{rastrigin [ms]}\\
+    \hline
+    10 \\
+    \hline
+
+    100 \\
+    \hline
+
+    1000 \\
+    \hline
+
+    10000 \\ 
+    \hline
+
+    100000 \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\noindent As we can see the execution time increases exponentially with the consequentially increase of
+the input.
+
+\section{Parallel implementation}
+
+\section{Parameters Retrievement}
+
+\section{Speedup and final conclusions}
+
 \end{multicols}

### 2025-02-11 code/bin_training_mpich.c
fix little problem

@@ -9,7 +9,8 @@
 #include <time.h>
 
 
-float eval(float *wi, unsigned int *seed, unsigned int d) {
+float eval(float *wi, unsigned int *seedi, unsigned int d) {
+  unsigned int seed = *seedi;
   for (int i = 0; i < 14; i++) {
     if (wi[i] < 0.0) {
       wi[i] = 0.0;
@@ -52,15 +53,15 @@ float eval(float *wi, unsigned int *seed, unsigned int d) {
 
     float *shifted_tmp = malloc(sizeof(float) * p.dim);
     float *shifted_rotation = malloc(sizeof(float) * p.dim * p.dim);
-    float *shifted_shift = init_array(p.dim, 80.0, seed);
-    init_matrix(shifted_rotation, 90.0, p.dim, seed);
+    float *shifted_shift = init_array(p.dim, 80.0, &seed);
+    init_matrix(shifted_rotation, 90.0, p.dim, &seed);
 
     init_shifted_fitness(shifted_tmp, shifted_rotation, shifted_shift,
                          rastrigin_fitness);
 
-    float *res = dragonfly_compute(p, w, fitness, 1, 0, 100.0, *seed);
+    float *res = dragonfly_compute(p, w, fitness, 1, 0, 100.0, seed);
     
-    avg += fitness(res, seed, p.dim);
+    avg += fitness(res, &seed, p.dim);
     /*printf("%f ", fitness(res, seed, p.dim));
     for(int i=0; i<14; i++){
       printf("%2f ", wi[i]);
@@ -88,7 +89,6 @@ int main(int argc, char **argv) {
   srand(time(NULL) + rank);
   Fitness fitness = eval;
   Parameters p = parameter_parse(argc, argv);
-  omp_set_num_threads(p.threads_per_process);
 
   float wi[14] = {
       0.000000,

### 2025-02-11 code/utils.c

@@ -191,7 +191,7 @@ void inner_dragonfly_step(Dragonfly *d, float *average_speed,
 
     // compute speed = sSi + aAi + cCi + fFi + eEi + w
 
-    for (unsigned int i = 0; i < dimensions; i++) {
+   for (unsigned int i = 0; i < dimensions; i++) {
       S = (cumulated_pos[i] / ((float)N)) - cur_pos[i];
       A = average_speed[i];
       C = (cumulated_pos[i] / (float)N) - cur_pos[i];
@@ -211,11 +211,12 @@ void inner_dragonfly_step(Dragonfly *d, float *average_speed,
     }
   }
 }
+
 void dragonfly_compute_step(Dragonfly *d, float *average_speed,
                             float *cumulated_pos, float *food, float *enemy,
                             unsigned int N, unsigned int NR_THREADS) {
   unsigned int base_random = rand_r(&d->seed);
-
+  //printf("base_random=%d\n", base_random);
   // if more than 1 thread
   unsigned int rest = d->N % NR_THREADS;
   unsigned int ratio = d->N / NR_THREADS;
@@ -334,9 +335,9 @@ fflush(stdout);
   }
   printf("END Function\n");
   fflush(stdout);
-}*/
+}
 
-void dragonfly_compute_step_serial(Dragonfly *d, float *average_speed,
+void dragonfly_compute_step(Dragonfly *d, float *average_speed,
                                    float *cumulated_pos, float *food,
                                    float *enemy, unsigned int N,
                                    unsigned int NUM_THREADS) {
@@ -391,7 +392,7 @@ void dragonfly_compute_step_serial(Dragonfly *d, float *average_speed,
 
   // update weights
   weights_step(&d->w);
-}
+}*/
 
 void computation_accumulate(ComputationStatus *status, Dragonfly *d,
                             float *best, float *best_fitness,

### 2025-02-13 text-paper/paper.tex
inserted openMP description and started MPI section into report

@@
 
-\section{Performace tests}
+\section{parallelization with MPI}
 
-The serial implementation seen above can be easily tried out using some test function 
-with known global minima in order to check the overall performances. We need to 
-report time execution for each test function, using the input dimension as independent variable;
-since however we have multiple input parameters to compare, we also have to decide which one will 
-be considered during the performance evaluation.
+OpenMPI is a library that enables code execution on multiple cores, wheter they belong
+to the same CPU or not; it can be useful in case of performace improvement of those algorihms
+that require to solve complex tasks, just like the dragonfly algorithm: in fact this library can be easily 
+adapted to the illustrated implementation in order to boost performance.
 
-\begin{table}[H]
-\begin{tabular}{| c | c | c | c |}
-  \hline
-  \textit{Nr of iterations} & \textbf{sphere [ms]} & \textbf{rosenblock [ms]} & \textbf{rastrigin [ms]}\\
-  \hline
-  10 \\
-  \hline
+\subsection*{MPI on chunks}
+As we said earlier chunks are independent from each otherm but usually they need to comunicate
+with their neighbours in order to keep updated the global status of the program execution. 
+It is possible to run chunks groups simoultaneously, using MPI builtin function to connect 
+the processes.  The instructions computed by each process are available in a custom function
+called \textit{dragonfly-compute}; after calculating the ID of the initial and final chunks
+and initializing message structures using the \textbf{offsetof} macro
+the function will execute the update loop that calculates the new velocities and positions 
+of the dragonfly for each chunk associated. To do this we first need to update the current
+location of enemies and food present in the chunk using the function \textbf{computation\_status\_merge},
+using the position of food and enemies of other chunks retrieved by the function \textbf{message\_broadcast};
+finally the process will proceed to update current position and velocities of dragonfiles using 
+the function \textbf{dragonfly\_compute\_step} illustrated earlier
 
-  100 \\
-  \hline
+\subsection*{MPI and parameters}
+Since MPI has been useful in the parallelization of the chunks then it can also become useful in
+the training of parameters
 
-  1000 \\
-  \hline
+\section{OpenMP implementation}
 
-  10000 \\ 
-  \hline
+OpenMP is a low level API that allows us to execute a sequence of instructions on different threads belonging
+to the same CPU core, improving in this way the performances of our implementation. This feature can be used to 
+speed up the execution of some slow functions, in our case these will be dragonfly-compute-step and message-accumulate
 
-  100000 \\
-  \hline
-\end{tabular}
+\subsection*{General use}
 
-\end{table}
+First of all it is necessary to understand how OpenMP should be inserted in the code: first of all we have to use 
+a compiler that supports this technology, in our case we have gcc, which allows OpenMP execution thanks to the flag -fopenmp.
+Second we need to include a pragma directive in our code, which will wrap the set of instructions that we are interested to execute
+in parallel. Since the implemented results will perform loops mainly then it is fundamental to calculate the work balance between the number of iterations, if the main
+body is a loop,
+and the number of threads available. Each thread has to define the starting and the end point of the loop that has to execute,
+calculating the starting index as $rank*ratio$ and the end one as $ratio*(rank+1)$, where $rank$ is the ID of the thread
 
-\begin{table}[H]
-  \begin{tabular}{| c | c | c | c |}
-    \hline
-    \textit{Nr of iterations} & \textbf{sphere [ms]} & \textbf{rosenblock [ms]} & \textbf{rastrigin [ms]}\\
-    \hline
-    10 \\
-    \hline
+\subsection*{dragonfly-compute-step}
 
-    100 \\
-    \hline
+One of the two functions that we are interested to execute using OpenMP is \textit{dragonfly-compute-step}:
+if we take a look on the serial version we notice that each computation step is independent
+from other iterations. This means that the execution flow can be split in many subtasks assigned
+to threads. The following code snippet illustrate how all the procedure has been implemented
+\begin{lstlisting}[language=C,caption={parallelized dragonfly-compute-step}]
 
-    1000 \\
-    \hline
+void dragonfly_compute_step(Dragonfly *d, 
+                            float *average_speed,
+                            float *cumulated_pos, 
+                            float *food, 
+                            float *enemy, 
+                            unsigned int N, 
+                            unsigned int NR_THREADS) {
 
-    10000 \\ 
-    \hline
+  unsigned int base_random = rand_r(&d->seed);
+  unsigned int rest = d->N % NR_THREADS;
+  unsigned int ratio = d->N / NR_THREADS;
 
-    100000 \\
-    \hline
-  \end{tabular}
-\end{table}
+  ...
 
-\noindent As we can see the execution time increases exponentially with the consequentially increase of
-the input.
+  #pragma omp parallel num_threads(NR_THREADS)
+    {
 
-\section{Parallel implementation}
+      unsigned int rank = omp_get_thread_num();
+      //printf("rank=%d\n", rank);
+      unsigned int base = ratio * rank;
+      unsigned int limit = ratio * (rank + 1);
+      inner_dragonfly_step(d, average_speed, 
+                          cumulated_pos, food, 
+                          enemy, N,
+                          base, limit, base_random);
+    }
 
-\section{Parameters Retrievement}
+  ...
 
-\section{Speedup and final conclusions}
+  if (rest != 0) {
+      unsigned int r_base = ratio * NR_THREADS;
+      unsigned int r_end = d->N;
+
+      inner_dragonfly_step(d, average_speed, 
+                           cumulated_pos, 
+                           food, enemy, N,
+                           r_base, r_end, base_random);
+    }
+  }
+
+  weights_step(&d->w);
+}
+
+\end{lstlisting}
+
+\noindent After we got all basic informations the \textit{inner\_dragonfly\_step}
+function can be executed, wrapping it into the \textbf{pragma} directive that we have seen earlier: in this way each thread will 
+perform the necessary operations only on a specific portion of the overall positions and speeds arrays.
+If the number of threads does not divide perfectly the size of the arrays then we compute the rest
+returned from the ratio calculation and the remaining cells will be processed using the serial approach.
+
+\subsection*{computation\_accumulate}
+
+The \textit{computation\_accumulate}  must compute
+the cumulated positions and speeds of the chunk and find the best food and enemy positions.
+The implementation of a parallelized approach in this case  is more complicated to achieve, 
+because threads should read and write on the same memeory areas which could be 
+difficult to execute independently from each other. This time it is necessary to define
+local buffers for each thread, so they can store temporary results without the risk of using
+overlapping memeory locations. The parallel section provides that each thread must compute the local
+cumulated sums of velocities and positions, then they must select the position with the best fitness value
+in order to find the best local position of food and enemies. Once the local execution is completed
+it will update the global status of the chunk using a critical section, with one thread that can execute that
+specific portion of the program
+
+\begin{figure}[H]
+  \includegraphics[scale=0.2]{reduce-example.png}
+  \centering
+  \caption{Illustration of the reduce process performed by the computation\_accumulate function}
+\end{figure}
 

### 2025-02-14 code/utils.c
fixed openMP accumulate implementation

@@ -394,37 +394,145 @@ void dragonfly_compute_step(Dragonfly *d, float *average_speed,
   weights_step(&d->w);
 }*/
 
+void inner_computation_accumulate(Dragonfly* d,float* cumulated_pos,float* cumulated_speed,float* next_food_fitness,
+																	float* next_enemy_fitness,float* best_fitness,
+																	unsigned int* indexes,unsigned int base,
+																	unsigned int limit,unsigned int dim,unsigned int seed){
+		
+		for (unsigned int k = base; k < limit; k++) {
+			float* iter_pos = d->positions + dim * k;
+			float* iter_speed = d->speeds + dim * k;
+			sum_assign(cumulated_pos,iter_pos,dim);
+			sum_assign(cumulated_speed,iter_speed,dim);
+						
+			float fitness = d->fitness(iter_pos, &seed, dim);
+
+			if (fitness > *next_food_fitness) {
+				indexes[0] = k;
+				*next_food_fitness = fitness;
+			}
+			if (fitness < *next_enemy_fitness) {
+				indexes[1] = k;
+				*next_enemy_fitness = fitness;
+			}
+			if (fitness > *best_fitness) {
+				indexes[2] = k;
+				*best_fitness = fitness;
+			}
+		}
+
+}
+
+void update_status(ComputationStatus* status,Dragonfly* d,float* best,float* best_fitness,unsigned int* indexes,unsigned int dim){
+		memcpy(status->next_food, d->positions + dim*indexes[0], sizeof(float) * dim);
+		memcpy(status->next_enemy, d->positions + dim*indexes[1], sizeof(float) * dim);
+		memcpy(best, d->positions + dim*indexes[2], sizeof(float) * dim);
+
+		status->next_food_fitness = d->fitness(status->next_food,&d->seed,dim);
+		status->next_enemy_fitness = d->fitness(status->next_enemy,&d->seed,dim);
+		*best_fitness = d->fitness(best,&d->seed,dim);
+}
+
 void computation_accumulate(ComputationStatus *status, Dragonfly *d,
                             float *best, float *best_fitness,
                             unsigned int NUM_THREADS) {
-  (void)(NUM_THREADS);
+
   unsigned int dim = d->dim;
-  zeroed(status->cumulated_pos, dim);
+	zeroed(status->cumulated_pos, dim);
   zeroed(status->cumulated_speeds, dim);
[…]

### 2025-02-15 text-paper/paper.tex
critique paragraph

@@
 
-\begin{multicols}{2}[\fontsize{9}{9}\section*{Abstract} \textbf{Dragonfly algorithm is a new kind of approach used to solve optimization problems through the emulation of dragonflies movements, but it lacks scalability.
+\begin{multicols}{2}[
+  \fontsize{9}{9}
+  \section*{Abstract}
+  \textbf{Dragonfly algorithm is a new kind of approach used to solve optimization problems through the emulation of dragonflies movements, but it lacks scalability.
 In this paper the authors try to set up an approach based on parallelization thanks to the use of API for high performance computing like MPI and OpenMP, then there will be a comparison of execution times between presented implementation and the classic approach based on serial programming, in order to see if parallelization could lead to a performance improvement and possible future extensions.}\newline]
@@
 
-Recent years have seen the raise and evolution of swarm intelligent algorithms, based on meta-heuristic approach where single individuals of a complex population follow an exploration-exploitation methodology to find the best solution in their local search space; each member of the group work independently from the others, leading to a subdivision of the labour which decrease the overall amount of execution time, working as decentralized and parallel units; but also, due to the fact that they mimic the behaviour of biological life forms, it is possible then that they could generate random patterns through the interaction with each other.
+Recent years have seen the raise and evolution of swarm intelligent algorithms, based on meta-heuristic approach where single individuals of a complex population follow an exploration-exploitation methodology to find the best solution in their local search space;
+each member of the group work independently from the others, leading to a subdivision of the labour which decrease the overall amount of execution time, working as decentralized and parallel units; but also, due to the fact that they mimic the behaviour of biological life forms, it is possible then that they could generate random patterns through the interaction with each other.
 The dragonfly algorithm has been proposed for the first time in 2016, and its purpose is to simulate the behaviour of a dragonfly population according to the principles of swarm intelligent algorihms seen previously.
@@
 \subsection*{How the algorithm works}
-The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$; then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program.
+The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$;
+then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program.
 This procedure is simulated through the observation of repetitive patterns observed in the behaviour of dragonflies, which are:
@@
   \item Attraction to a common source food, expressed by the formula $ F_i = F^+ - P$, where $F^+$ is the position of the current prey and $P$ is the position of the current dragonfly
-  \item Distraction of enemies, represented by the equation $ E_i = E^- - P$ where $E^-$ is the position of the current enemy 
+  \item Distraction of enemies, represented by the equation $ E_i = E^- + P$ where $E^-$ is the position of the current enemy 
 \end{itemize}
@@
 
-\noindent After the initialization of velocities and positions of agents,
+\noindent  As you can see this implementation use heavily the sum\_assign and similar functions, which we define to ease the computation and keep them similar to the mathematical notation.
+
+After the initialization of velocities and positions of agents,
 which will be represented by two separated matricies with the same dimension equal to the number of members of our
@@
   \item calculate product between weigths and arrays
-  \item update speed and position the dragonfly considering also levy flight
+  \item update speed and position the dragonfly considering also brownian motion
 \end{itemize}
+\section{Critical analysis of the coefficients}
+From this implementation we discovered some critical points that should be improved:
+\begin{itemize}
+  \item Chunks: In the paper and in the algorithm often uses the therm neighbours without giving a precise definition.
+  Assuming a general distance metric (eg Euclidean Distance), could be an option, but later in the paper it describes how the alghorithm runs in linear time with respect to the number of dragonflies and the iteration count.
+  This would be false in case of euclidean distance, because in order to define for each butterfly the neighbours, we should calculate the distance between each pair of dragonflies, which would lead to a time complexity of $O(N^2)$.
+  Maybe using some complex geospatial algorithms it could be possible to archive that, but we doubt this was the original intention of the authors.
+  Later in the paper also another crithical problem in respect to other algorithm was the premature convergence.
+So we decided to divide the population in chunks, and progressively merge chunks until we will remain with only one.
+  So in this way we addressed both problems.
+  \item Stability: In order to be a solid algorithm it must be garanteed that does not exist some values that would make the solution diverge.
+  As a first approach we introduced a max_speed parameter that would change over time and would assure stability.
+But given that in the original paper it was not described we omitted from later implementations.
+  This lead to instabilities from the enemy coefficient (if bigger than 0.5), but mainly from the separation coefficient.
+  This is a problem, because we could be see the coefficient as n times the distance between the dragonfly and the average position of the population.
+For example if the distance is 0.001, but there are 1000 dragonflies in the neighbourhood, then the coefficient would be 1, at the very next iteration the dragonflies would have a distance of 1 from the center and so on, causing an exponential growth.
+  In order to fix this we modify the separation formula as: $S_i = - \frac{\sum^M_{j=1}{P - P_j}}{N}$.
+  \item  The enemy coefficient does not represent what the author describes. The sum of vectors does not necessarily point outwards in respect to the enemy (eg with aligned pos vectors). 
+  We assumed a typo, but even the difference of the vectors would not be enough, because if the enemy is too close to the dragonfly, the difference would get smaller and not bigger.
+  A possible solution would be to normalize the difference vector with the inverse of the vector length, but the formulation would have been to different from the original one.
+So we decided to keep the original formulation and see if it actualy would do something.
+
 \subsection*{chunks}
+
 We can improve this approach introducing groupments of dragonflies into chunks, 
@@
 \end{figure}
+
+A
 

### 2025-02-15 text-paper/paper.tex
started results section in the report

@@
 \usepackage{longtable}
+\usepackage{cite}
 
@@
 \noindent Once all the parameters have been calculated we can now update the current step vector of each dragonfly using the following equation: $$\Delta P^{t+1}_{i} = (sS_i + aA_i + fF_i + cC_i) + \omega\Delta P^t_i$$ where $s,a,f,c$ are respectevly the separation weight, the alignment weight, the attraction weight, the cohesion weight and $\omega$ is the inertia weight, all of them are randomly chosen in the interval $[0,1]$.
-When all step vectors have been updated we can now set the new position of the dragonflies using this equation, which takes the new step vector as an input: $$P^{t+1}_{i} = P_i^t + \Delta P_{i}^{t+1}$$.
-However if there is a dragonfly which does not have any neighbour then we assume that the overall population moves in a random way: the most common method to do that is to use a random walk function.
+When all step vectors have been updated we can now set the new position of the dragonflies using this equation, which takes the new step vector as an input: $$P^{t+1}_{i} = P_i^t + \Delta P_{i}^{t+1}$$However if there is a dragonfly which does not have any neighbour then we assume that the overall population moves in a random way: the most common method to do that is to use a random walk function.
 The previous formula then must be modified into this one here: $$P^{t+1}_i = P_i^t + Levy(d)$$ where $d$ is the dimension of the position vector, and $Levy(d)$ is the Levy flight function, which is calculated in this way: $$Levy(d) = 0.01\times\frac{r_1 \times \sigma}{|r_2|^{\frac{1}{\beta}}}$$ where $r1$ and $r2$ are two random numbers uniformally distributed in the interval $[0,1]$, and $\sigma$ is a vector calculated using the following formula: $$\sigma = \frac{sin(\frac{\beta\pi}{2})\times\Gamma(1+\beta)}{\Gamma(\frac{\beta+1}{2})\times \beta \times 2^{(\beta-1)/2}}$$ where $\Gamma(x)$ is the gamma function \cite{WikiGamma}, and $\beta$ is a constant.Another valid approach is to apply, instead of Levy flight, the brownian motion model, used especially in gas and particle analysis: this seems to return better performances \cite{BDragonfly}.
 
-\section{first approach}
+\section{General approach}
 
@@
                         float *cumulated_pos, 
-                          float * food, 
-                          float * enemy, 
-                          unsigned int N) {
+                        float *food, 
+                        float *enemy,
+                        unsigned int N, 
+                        unsigned int random_seed) {
 
-    unsigned int dim = d->dim;
-      // for each dragonfly
+  unsigned int dimensions = d->dim;
+  float S;
+  float A;
+  float C;
+  float F;
+  float E;
+  float levy;
+
+  float *cur_pos;
+  float *cur_speed;
+
   for (unsigned int j = 0; j < d->N; j++) {
-      float *cur_pos = d->positions + dim * j;
-      float *cur_speed = d->speeds + dim * j;
-
-      // compute separation: Si = -sumall(X-Xi)
-      memcpy(d->S, cumulated_pos, sizeof(float) * dim);
-      scalar_prod_assign(d->S, 1.0/(float)N, dim);
-      sub_assign(d->S, cur_pos, dim);
-      scalar_prod_assign(d->S, d->w.s, dim);
-
-      // compute alignament: Ai = avarage(Vi)
-      memcpy(d->A, average_speed, sizeof(float) * dim);
-      scalar_prod_assign(d->A, d->w.a, dim);
-
-      // compute cohesion: Ci = avarage(Xi)-X
-      memcpy(d->C, cumulated_pos, sizeof(float) * dim);
-      scalar_prod_assign(d->C, 1.0 / (float)N, dim);
-      sub_assign(d->C, cur_pos, dim);
-      scalar_prod_assign(d->C, d->w.c, dim);
-
-      // food attraction: Fi=X_food - X
-      memcpy(d->F, food, sizeof(float) * dim);
-      sub_assign(d->F, cur_pos, dim);
-      scalar_prod_assign(d->F, d->w.f, dim);
-
-      // enemy repulsion: E=X_enemy+X
-      memcpy(d->E, enemy, sizeof(float) * dim);
-      sum_assign(d->E, cur_pos, dim);
-      scalar_prod_assign(d->E, d->w.e, dim);
-
-      brownian_motion(d->levy, dim, &d->seed);
+    unsigned random = random_seed + j;
+    cur_pos = d->positions + j * dimensions;
+    cur_speed = d->speeds + j * dimensions;
 
     // compute speed = sSi + aAi + cCi + fFi + eEi + w
-      scalar_prod_assign(cur_speed, d->w.w, dim);
-      sum_assign(cur_speed, d->E, dim);
-      sum_assign(cur_speed, d->F, dim);
-      sum_assign(cur_speed, d->C, dim);
-      sum_assign(cur_speed, d->A, dim);
-      sum_assign(cur_speed, d->S, dim);
-      sum_assign(cur_speed, d->levy, dim);
 
-      // update current pos
-      sum_assign(cur_pos, cur_speed, dim);
+   for (unsigned int i = 0; i < dimensions; i++) {
+      S = ((cumulated_pos[i] - cur_pos[i]) / (float)N);
+      A = average_speed[i];
+      C = (cumulated_pos[i] / (float)N) - cur_pos[i];
+      F = food[i] - cur_pos[i];
+      E = enemy[i] + cur_pos[i];
+      levy = RAND_FLOAT(1.0, &random);
+
+      cur_speed[i] *= d->w.w;
+      cur_speed[i] += d->w.s * S;
+      cur_speed[i] += d->w.a * A;
+      cur_speed[i] += d->w.c * C;
+      cur_speed[i] += d->w.f * F;
+      cur_speed[i] += d->w.e * E;
+      cur_speed[i] += levy;
+
+      cur_pos[i] += cur_speed[i];
     }
-
-    // update weights
-    weights_step(&d->w);
+  }
 }
@@
 
-\noindent  As you can see this implementation use heavily the sum\_assign and similar functions, which we define to ease the computation and keep them similar to the mathematical notation.
+\noindent After the initialization of velocities and positions of agents,
+which will be represented by two separated matricies with the same dimension equal to the number of members of our
+population times the dimension of our problem, the algorithm starts to execute a for loop 
+with a fixed amount of iteration stored in the variable $d\rightarrow N$,  retrieve
+the position and the velocity of the current dragonfly, represented by pointers
+\textit{cur\_pos} and \textit{cur\_speed},
+and the cumulated coordinates of its neighbours. 
+The execution then proceed with the following steps:
 
-After the initialization of velocities and positions of agents,
-which will be represented by two separated matricies with the same dimension equal to the number of members of our
-population times the dimension of our problem, the algorithm starts to execute a for loop with a fixed amount of iteration stored in the variable $d\rightarrow N$, 
-and will repeatetly compute the following steps:
 \begin{itemize}
@@
   \item compute separation, alignment, cohesion, food attraction and enemy distraction vectors,
-  stored into arrays placed in the struct Dragonfly
-  \item calculate product between weigths and arrays
+  stored into local variables named with their abrreviations
+  \item calculate product between weigths and and the single coordinates of the velocity array
   \item update speed and position the dragonfly considering also brownian motion
 \end{itemize}
+
+\noindent Notice how the computation
+of velocities and speed coordinates is compacted into the same \textbf{for} loop;
+this allows to reduce time execution, because the overall calculation is made locally
+all in the same place, leaving out the necessity of using auxiliary arrays for the coefficients.
+The reason why we divide the separation coefficient for the number of neighbours 
+will be explained later in this paper
+
+\subsection*{chunks}
+
+As anticipated earlier dragonfly algorithm, as many other swarm optimization solutions,
+has a good chance to be stuck into a local minima, without the possibility to find better
+candidates. We can avoid this and improve this approach introducing groupments of dragonflies into chunks, 
+calculating local minima of different areas and compare them in order
+to get as close as possible to the global solution.
+
+\begin{figure}[H]
+  \includegraphics[scale=0.3]{img/chunks.png}
+  \centering
+  \caption{illustration of communication between chunks}
+\end{figure}
+
+\noindent The implementation of this process is quite simple: each chunk will contain a subset
+of components of the global population that will specialize to compute the point with the best local
+fitness for both enemies and food of that specific region, then the local status will be communicated
+to all other chunks, updating in this way the global status of the program thanks to the update of positions,
+and the local configuration of each region; the algorithm ends displaying the best fitness found
+during the execution.
+
 \section{Critical analysis of the coefficients}
@@
 \begin{itemize}
-  \item Chunks: In the paper and in the algorithm often uses the therm neighbours without giving a precise definition.
+  \item In the original paper\cite{Original} and in the algorithm often uses the therm neighbours without giving a precise definition.
   Assuming a general distance metric (eg Euclidean Distance), could be an option, but later in the paper it describes how the alghorithm runs in linear time with respect to the number of dragonflies and the iteration count.
@@
   So in this way we addressed both problems.
-  \item Stability: In order to be a solid algorithm it must be garanteed that does not exist some values that would make the solution diverge.
-  As a first approach we introduced a max_speed parameter that would change over time and would assure stability.
+  \item In order to be a solid algorithm it must be garanteed that does not exist some values that would make the solution diverge.
+  As a first approach we introduced a max\_speed\ parameter that would change over time and would assure stability. 
   But given that in the original paper it was not described we omitted from later implementations.
   This lead to instabilities from the enemy coefficient (if bigger than 0.5), but mainly from the separation coefficient.
-  This is a problem, because we could be see the coefficient as n times the distance between the dragonfly and the average position of the population.
-For example if the distance is 0.001, but there are 1000 dragonflies in the neighbourhood, then the coefficient would be 1, at the very next iteration the dragonflies would have a distance of 1 from the center and so on, causing an exponential growth.
-  In order to fix this we modify the separation formula as: $S_i = - \frac{\sum^M_{j=1}{P - P_j}}{N}$.
+  This is a problem, because we could be see the coefficient as n times the distance between the dragonfly and the average position of the population. For example if the distance is 0.001, 
+  but there are 1000 dragonflies in the neighbourhood, then the coefficient would be 1, at the very next iteration the dragonflies would have a distance of 1 from the center and so on, causing an exponential growth.
+  In order to fix this we modify the separation formula as: $S_i = - \frac{\sum^M_{j=1}{P - P_j}}{M}$.
+
   \item  The enemy coefficient does not represent what the author describes. The sum of vectors does not necessarily point outwards in respect to the enemy (eg with aligned pos vectors). 
@@
 So we decided to keep the original formulation and see if it actualy would do something.
+\end{itemize}
 
-\subsection*{chunks}
-
-We can improve this approach introducing groupments of dragonflies into chunks, 
-allowing us to calculate local minima of different areas and compare them in order
-to get as close as possible to the global minima. which will be considered
-autonomous agents that will calculate local minima of our fitness function, updating then the 
-global status of by sending messages to all other dragonfly chunks.
-
-\begin{figure}[H]
-  \includegraphics[scale=0.3]{chunks.png}
-  \caption{illustration of communication between chunks}
-\end{figure}
-
-A
-
-\section{parallelization with MPI}
+\section{Parallelization with MPI}
 
@@
 Since MPI has been useful in the parallelization of the chunks then it can also become useful in
-the training of parameters
+the training of parameters.
 
@@
 to threads. The following code snippet illustrate how all the procedure has been implemented
+
 \begin{lstlisting}[language=C,caption={parallelized dragonfly-compute-step}]
-
 void dragonfly_compute_step(Dragonfly *d, 
@@
   weights_step(&d->w);
+
   }
@@
 local buffers for each thread, so they can store temporary results without the risk of using
-overlapping memeory locations. The parallel section provides that each thread must compute the local
+overlapping memeory locations. 
+\begin{figure}[H]
+  \includegraphics[scale=0.2]{img/reduce-example.png}
+  \centering
+  \caption{Illustration of the reduce process performed by the computation\_accumulate function}
+\end{figure}
+\noindent The parallel section provides that each thread must compute the local
 cumulated sums of velocities and positions, then they must select the position with the best fitness value
@@
 
+\section{Final execution and results}
+Now that we have described all the procedure it is time to illustrate how 
+the algorithm has been executed. This job has been performed by the
+High Performance Computing cluster at University of Trento, which 
+use a particular technology called PBS to handle multiple tasks at the same
+time: this feature comes useful in our case since, as we will see later, we 
+execute multiple istances of the same problem
+\subsection*{structure of the execution}
+
+As anticipated we will execute the same problem using multiple instances with different amount of
+resources in order to get a complete evaluation of performance. The overall execution use from 1 to 8
+threads, and 4 to 1024 cores
+
+\subsection*{performace theory}
+
+After the end of all tasks it is possible to check how much gain in performance we have obtained.
+To do this it is necessary to introduce some theoretical concepts that will give us the tools that
+will give us a good metric system to evaluate our results.Speedup is defined as the ratio between the the execution times of the serial and parallel
+implementations, so the formula will be equal to: $$Speedup = \frac{T_{serial}}{T_{parallel}}$$ 
+from this definition it is obvious that the more the speedup is the more the parallel implementation
+performs better. There is a particular case where the time execution of the denominator can be defined
+as $T_{parallel}=T_{serial}/p$, where \textit{p} is the number of cores: this is called \textit{linear speedup},
+and in this situation it perfectly match with the classic speedup coefficient
+Efficiency is another important parameter, and it is defined by the ratio of the speedup and the 
+number of cores, giving the following formula: $$E = \frac{Speedup}{p}$$
+also in this case the efficiency decrease due to the increase of the number of cores. In case
+of linear speedup we get that the efficiency is equal to the inverse of $p$
+\subsection*{results}
+Now it is possible to show results obtained from the cluster,
+using the configuration with 4 cores and 1 thread will be considered
+equal as the serial time
+\newline
+
+\begin{tabular}{| c | c | c | c | c | c | c |}
+  \hline
+  T/p & 4	 & 8	& 16	& 32	& 64 & 128  \\
+  \hline
+  1 & 1	& 2,03 & 3,87 & 7,94 & 14,76 & 31,19  \\
+  2 & 1,04 & 2,05 & 3,97 & 8,50 & 16,43 & 27,90 \\	
+  4 & 1,01 &2,03	& 4,21 &	8,04 &	16,37 &	35,26	\\	
+  8 & 1,02  &2,02 &	4,02 &	8,62 &	18,32 &	32,26 \\
+  \hline
+\end{tabular}
+\newline\newline As we can see the speedup tends to 
+increase when the number of cores increase, however when
+we use more threads it happens that the
+speedup start to decrease at a certain point for some
+core configuration. As we can see from the plots it 
+seems there are threshold at some specific points, depending
+on the number of threads used; notice how the inclinations
+of the lines in the plot tends to be less and less evident.
+
 \begin{figure}[H]
-  \includegraphics[scale=0.2]{reduce-example.png}
+  \includegraphics[scale=0.3]{img/speedup.jpg}
   \centering
-  \caption{Illustration of the reduce process performed by the computation\_accumulate function}
+  \caption{speedup graphic}
+\end{figure}
+
+\noindent 
+Efficiency tends to present an unpredictable behaviour,
+in fact the coefficients grows for some particular setups,
+for others instead numbers decrease. If we take a look on
+the plot below it shows that there is a high variance in the
+datas, with the lines that tends to be more stretched when the
+number of threads gets higher
+\newline
+\newline
+\begin{tabular}{| c | c | c | c | c | c | c |}
+  \hline
+  T/p & 4	 & 8	& 16	& 32	& 64 & 128  \\
+  \hline
+  1 & 1	& 1,02 & 0,97 & 0,99 & 0,92 & 0,97  \\
+  2 & 1,04 & 1,03 & 0,99 & 1,06 & 1,03 & 0,87 \\	
+  4 & 1,01	&1,01&	1,05&	1&	1,02&	1,1	\\	
+  8 &1,02&	1,01&	1,01&	1,08&	1,15&	1,01 \\
+  \hline
+\end{tabular}
+
+\begin{figure}[H]
+  \includegraphics[scale=0.3]{img/efficiency.jpg}
+  \centering
+  \caption{efficiency graphic}
 \end{figure}

### 2025-08-21 code/bin_mpich.c
it works and computes + fmt

@@ -10,7 +10,6 @@
 #include "dragonfly-common.h"
 #include "utils.h"
 
-
 int main(int argc, char *argv[]) {
   MPI_Init(NULL, NULL);
   // wait for all the process to start
@@ -41,7 +40,7 @@ int main(int argc, char *argv[]) {
   float *res = dragonfly_compute(p, w, fitness, comm_size, rank, 100.0, 0);
 
   MPI_Barrier(MPI_COMM_WORLD);
-  unsigned int s =0;
+  unsigned int s = 0;
   float fit = fitness(res, &s, p.problem_dimensions);
 
   if (rank == 0) {

### 2025-08-21 code/bin_mpich_2.c

@@ -10,13 +10,10 @@
 #include "dragonfly-common.h"
 #include "utils.h"
 
-
 int main(int argc, char *argv[]) {
   MPI_Init(NULL, NULL);
   // wait for all the process to start
   MPI_Barrier(MPI_COMM_WORLD);
-
-  
 
   // start clock
   clock_t start_time;
@@ -36,7 +33,7 @@ int main(int argc, char *argv[]) {
       .wl = {0.6, 0.6},
       .ll = {0.1, 0.1},
 
-      //chuk size
+      // chuk size
   };
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
 
@@ -48,7 +45,7 @@ int main(int argc, char *argv[]) {
   float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 100.0, 0);
 
   MPI_Barrier(MPI_COMM_WORLD);
-  unsigned int s =0;
+  unsigned int s = 0;
   float fit = fitness(res, &s, p.problem_dimensions);
 
   if (rank == 0) {

### 2025-08-21 code/bin_serial.c

@@ -16,7 +16,8 @@ int main(int argc, char *argv[]) {
   Parameters p = parameter_parse(argc, argv);
   unsigned int seed = time(NULL);
   float *shifted_tmp = malloc(sizeof(float) * p.problem_dimensions);
-  float *shifted_rotation = malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
+  float *shifted_rotation =
+      malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
   float *shifted_shift = init_array(p.problem_dimensions, 100.0, &seed);
   init_matrix(shifted_rotation, 100.0, p.problem_dimensions, &seed);
 
@@ -24,20 +25,9 @@ int main(int argc, char *argv[]) {
                        rastrigin_fitness);
 
   Fitness fitness = shifted_fitness;
-  float wi[14] = {0.803008,
-0.801540,
-0.000000,
-4.131542,
-1.180980,
-2.213441,
-1.289022,
-0.828981,
-0.100000,
-0.022841,
-0.000000,
-0.000000,
-0.108948,
-0.404676};
+  float wi[14] = {0.803008, 0.801540, 0.000000, 4.131542, 1.180980,
+                  2.213441, 1.289022, 0.828981, 0.100000, 0.022841,
+                  0.000000, 0.000000, 0.108948, 0.404676};
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},
@@ -51,7 +41,7 @@ int main(int argc, char *argv[]) {
   };
 
   float *res = dragonfly_compute(p, w, fitness, 1, 0, 100.0, seed);
-  unsigned int s =0;
+  unsigned int s = 0;
   float fit = fitness(res, &s, p.problem_dimensions);
 
   printf("found fitness=%f\n", fit);

### 2025-08-21 code/bin_training_mpich.c

@@ -8,31 +8,31 @@
 #include <string.h>
 #include <time.h>
 
-
 float eval(float *wi, unsigned int *seedi, unsigned int d) {
   unsigned int seed = *seedi;
   for (int i = 0; i < 14; i++) {
     if (wi[i] < 0.0) {
       wi[i] = 0.0;
     }
-    if(wi[i]>2.0){
-      wi[i]=2.0;
+    if (wi[i] > 2.0) {
+      wi[i] = 2.0;
     }
     if (i % 2 == 0 && wi[i] == wi[i + 1]) {
       wi[i + 1] += 0.0001;
     }
   }
   float m = max(wi[8], wi[9]);
-  if(m>0.1){
-    wi[8]/=10*m;
-    wi[9]/=10*m;
+  if (m > 0.1) {
+    wi[8] /= 10 * m;
+    wi[9] /= 10 * m;
   }
   m = max(wi[10], wi[11]);
-  if(m>1.0){
-    wi[10]/=1.2*m;
-    wi[11]/=1.2*m;
+  if (m > 1.0) {
+    wi[10] /= 1.2 * m;
+    wi[11] /= 1.2 * m;
   }
   (void)d;
+
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},
@@ -44,23 +44,28 @@ float eval(float *wi, unsigned int *seedi, unsigned int d) {
       .wl = {wi[10], wi[11]},
       .ll = {wi[12], wi[13]},
   };
-  Parameters p = {.population_size = 128, .problem_dimensions = 10, .starting_chunk_count = 64, .iterations = 80, .threads_per_process = 1};
-
+  Parameters p = {.population_size = 128,
+                  .problem_dimensions = 10,
+                  .starting_chunk_count = 64,
+                  .iterations = 80,
+                  .threads_per_process = 1};
+  ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
   Fitness fitness = shifted_fitness;
   float avg = 0.0;
   int N = 30;
   for (int i = 0; i < N; i++) {
 
     float *shifted_tmp = malloc(sizeof(float) * p.problem_dimensions);
-    float *shifted_rotation = malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
+    float *shifted_rotation =
+        malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
[…]

### 2025-08-21 code/bin_training_parameters.c

@@ -1,35 +1,33 @@
 #include "dragonfly-common.h"
 #include "utils.h"
+#include <math.h>
 #include <stdbool.h>
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
-#include<math.h>
 #include <time.h>
-
-
 
 float eval(float *wi, unsigned int *seed, unsigned int d) {
   for (int i = 0; i < 14; i++) {
     if (wi[i] < 0.0) {
       wi[i] = 0.0;
     }
-    if(wi[i]>2.0){
-      wi[i]=2.0;
+    if (wi[i] > 2.0) {
+      wi[i] = 2.0;
     }
     if (i % 2 == 0 && wi[i] == wi[i + 1]) {
       wi[i + 1] += 0.0001;
     }
   }
   float m = max(wi[8], wi[9]);
-  if(m>0.1){
-    wi[8]/=10*m;
-    wi[9]/=10*m;
+  if (m > 0.1) {
+    wi[8] /= 10 * m;
+    wi[9] /= 10 * m;
   }
   m = max(wi[10], wi[11]);
-  if(m>1.0){
-    wi[10]/=1.2*m;
-    wi[11]/=1.2*m;
+  if (m > 1.0) {
+    wi[10] /= 1.2 * m;
+    wi[11] /= 1.2 * m;
   }
   (void)d;
   Weights w = {
@@ -43,37 +41,41 @@ float eval(float *wi, unsigned int *seed, unsigned int d) {
       .wl = {wi[10], wi[11]},
       .ll = {wi[12], wi[13]},
   };
-  Parameters p = {.population_size = 128, .problem_dimensions = 10, .starting_chunk_count = 64, .iterations = 200};
+  Parameters p = {.population_size = 128,
+                  .problem_dimensions = 10,
+                  .starting_chunk_count = 64,
+                  .iterations = 200};
 
   Fitness fitness = shifted_fitness;
   int N = 30;
-  float *array = malloc(sizeof(float)*N);
+  float *array = malloc(sizeof(float) * N);
   for (int i = 0; i < N; i++) {
 
     float *shifted_tmp = malloc(sizeof(float) * p.problem_dimensions);
-    float *shifted_rotation = malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
+    float *shifted_rotation =
+        malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
[…]

### 2025-08-21 code/dragonfly-common.c

@@ -218,9 +218,14 @@ unsigned int assing_logical_chunks(unsigned int phisical_chunk_start,
   unsigned int index = 0;
   while (cur_start < phisical_chunk_end) {
     unsigned int cur_end = min(cur_start + logical_chunk_size, population_size);
-    bool starts_in_previous = cur_start<phisical_chunk_start && phisical_chunk_start<cur_end && cur_end<phisical_chunk_end;
-    unsigned int next_phisical_end = remaining<=1 ? population_size: phisical_chunk_end+phisical_chunk_size;  
-    bool ends_in_next = phisical_chunk_end<cur_end && cur_end<next_phisical_end;
+    bool starts_in_previous = cur_start < phisical_chunk_start &&
+                              phisical_chunk_start < cur_end &&
+                              cur_end < phisical_chunk_end;
+    unsigned int next_phisical_end =
+        remaining <= 1 ? population_size
+                       : phisical_chunk_end + phisical_chunk_size;
+    bool ends_in_next =
+        phisical_chunk_end < cur_end && cur_end < next_phisical_end;
     LogicalChunk cur = {
         .start = cur_start,
         .end = cur_end,
@@ -228,9 +233,9 @@ unsigned int assing_logical_chunks(unsigned int phisical_chunk_start,
         .to_shift_right = !starts_in_previous && ends_in_next,
         .complete = !starts_in_previous && !ends_in_next,
     };
-    //printf("chunk: %d-%d %d-%d %d %d %d\n", phisical_chunk_start,
-    //         phisical_chunk_end, cur_start, cur_end, cur.to_shift_left,
-    //         cur.to_shift_right, index);
+    // printf("chunk: %d-%d %d-%d %d %d %d\n", phisical_chunk_start,
+    //          phisical_chunk_end, cur_start, cur_end, cur.to_shift_left,
+    //          cur.to_shift_right, index);
     buffer[index] = cur;
     index++;
     cur_start += logical_chunk_size;
@@ -238,60 +243,67 @@ unsigned int assing_logical_chunks(unsigned int phisical_chunk_start,
   return index;
 }
 unsigned int upper_log2(unsigned int n) {
-    unsigned int log = 0;
-    if (n == 0) return 0; // or handle as needed
-    n--;
[…]

### 2025-08-21 code/dragonfly-common.h

@@ -4,39 +4,39 @@
 #define MESSAGE_SIZE 16
 #define MAX_CHUNKS 4048
 
-
-// struct used to store the progression of the dragonfly weights s, a, c, f, e, w, l (linear)
+// struct used to store the progression of the dragonfly weights s, a, c, f, e,
+// w, l (linear)
 
 typedef struct {
   float sl[2], al[2], cl[2], fl[2], el[2], wl[2], ll[2];
   float st, at, ct, ft, et, wt, lt;
   float s, a, c, f, e, w, l;
 
-
 } Weights;
 
 // also it stores the logical_chunk_count progression (geometric)
 typedef struct {
-    unsigned start_count, end_count;
+  unsigned start_count, end_count;
   unsigned int current_step;
   unsigned int total_steps;
   unsigned count;
 
 } ChunkSize;
 
-ChunkSize new_chunk_size(unsigned int start_count, unsigned int end_count, unsigned int steps);
+ChunkSize new_chunk_size(unsigned int start_count, unsigned int end_count,
+                         unsigned int steps);
 
-void update_chunk_size(ChunkSize* c);
+void update_chunk_size(ChunkSize *c);
 
-// Newtype for Fitness functions, they should take the input vector, one int as random number generator seed, and the dimension count
-typedef float (*Fitness)(float *, unsigned int*, unsigned int);
+// Newtype for Fitness functions, they should take the input vector, one int as
+// random number generator seed, and the dimension count
+typedef float (*Fitness)(float *, unsigned int *, unsigned int);
 
 // function used to compute the step for the linear progression of the weights.
 void weights_compute_steps(Weights *w, unsigned int steps);
 
 // compute one step forward for the weights
 void weights_step(Weights *w);
-
 
 typedef struct {
   // dimensions of the problem
@@ -45,32 +45,29 @@ typedef struct {
   float space_size;
   // fitness function
   Fitness fitness;
-  //adaptive weights
+  // adaptive weights
   Weights w;
-  
+
   // start and end of elements used by the current thread
   unsigned int start, end;
-  
+
   // local_buffers the size is local_end-local_start elements
   float *positions, *speeds;
 
   // tmp buffers (in order to not allocate and deallocate memory)
   float *S, *A, *C, *F, *E, *W, *delta_pos, *levy;
-  
-  //random seed
+
+  // random seed
   unsigned int seed;
 } Dragonfly;
 
+Dragonfly dragonfly_new(unsigned int dimensions, unsigned int start,
[…]

### 2025-08-21 code/utils.c

@@ -170,7 +170,7 @@ float shifted_fitness(float *inp, unsigned int *seed, unsigned int dim) {
 #include <omp.h>
 
 #endif
-
+/*
 void inner_dragonfly_step(Dragonfly *d, float *average_speed,
                           float *cumulated_pos, float *food, float *enemy,
                           unsigned int N, unsigned int base, unsigned int limit,
@@ -213,7 +213,7 @@ void inner_dragonfly_step(Dragonfly *d, float *average_speed,
     }
   }
 }
-/*
+
 void dragonfly_compute_step(Dragonfly *d, float *average_speed,
                             float *cumulated_pos, float *food, float *enemy,
                             unsigned int N, unsigned int NR_THREADS) {

### 2025-08-22 code/bin_training_mpich.c
works

@@ -5,8 +5,29 @@
 #include <stdbool.h>
 #include <stdio.h>
 #include <stdlib.h>
-#include <string.h>
 #include <time.h>
+#include <unistd.h>
+#include<errno.h>
+int msleep(long msec)
+{
+    struct timespec ts;
+    int res;
+
+    if (msec < 0)
+    {
+        errno = EINVAL;
+        return -1;
+    }
+
+    ts.tv_sec = msec / 1000;
+    ts.tv_nsec = (msec % 1000) * 1000000;
+
+    do {
+        res = nanosleep(&ts, &ts);
+    } while (res && errno == EINTR);
+
+    return res;
+}
 
 float eval(float *wi, unsigned int *seedi, unsigned int d) {
   unsigned int seed = *seedi;
@@ -32,7 +53,6 @@ float eval(float *wi, unsigned int *seedi, unsigned int d) {
     wi[11] /= 1.2 * m;
   }
   (void)d;
-
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},
@@ -77,7 +97,6 @@ float eval(float *wi, unsigned int *seedi, unsigned int d) {
     free(shifted_shift);
     free(res);
   }
-
   return avg / N;
 }
 
@@ -112,7 +131,7 @@ int main(int argc, char **argv) {
       .ll = {wi[12], wi[13]},
   };
   float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 2.0, 0);
-
+  //float *res = malloc(sizeof(float)*p.population_size); 
   MPI_Barrier(MPI_COMM_WORLD);
 
   if (rank == 0) {

### 2025-08-22 code/dragonfly-common.c

@@ -362,7 +362,7 @@ void inner_dragonfly_step(Dragonfly *d, LogicalChunk chunk,
                           unsigned int random_seed) {
 
   unsigned int dimensions = d->dim;
-  printf("%d\n", dimensions);
+  //printf("%d\n", dimensions);
   float S;
   float A;
   float C;
@@ -409,6 +409,7 @@ void inner_dragonfly_step(Dragonfly *d, LogicalChunk chunk,
 float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
                          unsigned int threads, unsigned int rank_id,
                          float space_size, unsigned int srand) {
+                          
   if (MESSAGE_SIZE < p.problem_dimensions) {
     fprintf(stderr,
             "impossible to compute with %d dimensions, recompile with a bigger "
@@ -456,10 +457,14 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
       (logical_span + logical_chunk_size - 1) / logical_chunk_size + 1;
 
   // define mpi structs (if available)
-  // #ifdef USE_MPI
-  MPI_Datatype computation_status_type;
-  Build_mpi_type_computation_status(&computation_status_type);
-  // #endif
+  void *computation_status_type=NULL;
+  if (threads > 1) {
+    computation_status_type=malloc(sizeof(MPI_Datatype));
+    //MPI_Datatype computation_status_type;
+     Build_mpi_type_computation_status(computation_status_type);
+  }
+  
+  // #endif// #ifdef USE_MPI
 
   LogicalChunk *local_chunks =
       malloc(sizeof(LogicalChunk) * maximum_logical_chunks);
@@ -473,12 +478,10 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
 
   dragonfly_alloc(&cur);
   ComputationStatus temp_comp;
-
   float *best = malloc(sizeof(float) * p.problem_dimensions);
   memcpy(best, cur.positions, sizeof(float) * p.problem_dimensions);
   unsigned seed = 0;
   float best_fitness = cur.fitness(best, &seed, p.problem_dimensions);
-
   // exit(0);
   //  MAIN COMPUTATION
   for (unsigned int i = 0; i < p.iterations; i++) {
[…]

### 2025-08-22 code/bin_mpich_2.c
better parameters

@@ -21,19 +21,32 @@ int main(int argc, char *argv[]) {
 
   // set parameters
   Parameters p = parameter_parse(argc, argv);
-
+float wi[] = {
+  0.0f,
+  2.0f,
+  0.0f,
+  0.0001f,
+  0.0f,
+  0.352596f,
+  2.0f,
+  2.0f,
+  0.0f,
+  0.0001f,
+  0.600357f,
+  0.0f,
+  0.0f,
+  0.0001f
+};
   Weights w = {
       // exploring
-      .al = {0.1, 0.1},
-      .cl = {0.7, 0.7},
+      .al = {wi[0], wi[1]},
+      .cl = {wi[2], wi[3]},
       // swarming
-      .sl = {0.1, 0.1},
-      .fl = {1.0, 1.0},
-      .el = {0.0, 0.0},
-      .wl = {0.6, 0.6},
-      .ll = {0.1, 0.1},
-
-      // chuk size
+      .sl = {wi[4], wi[5]},
+      .fl = {wi[6], wi[7]},
+      .el = {wi[8], wi[9]},
+      .wl = {wi[10], wi[11]},
+      .ll = {wi[12], wi[13]},
   };
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
 
@@ -42,7 +55,7 @@ int main(int argc, char *argv[]) {
   MPI_Comm_rank(MPI_COMM_WORLD, &rank);
   Fitness fitness = rastrigin_fitness;
 
-  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 100.0, 0);
+  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 100.0, start_time);
 
   MPI_Barrier(MPI_COMM_WORLD);
   unsigned int s = 0;

### 2025-08-22 code/bin_training_mpich.c

@@ -7,27 +7,6 @@
 #include <stdlib.h>
 #include <time.h>
 #include <unistd.h>
-#include<errno.h>
-int msleep(long msec)
-{
-    struct timespec ts;
-    int res;
-
-    if (msec < 0)
-    {
-        errno = EINVAL;
-        return -1;
-    }
-
-    ts.tv_sec = msec / 1000;
-    ts.tv_nsec = (msec % 1000) * 1000000;
-
-    do {
-        res = nanosleep(&ts, &ts);
-    } while (res && errno == EINTR);
-
-    return res;
-}
 
 float eval(float *wi, unsigned int *seedi, unsigned int d) {
   unsigned int seed = *seedi;
@@ -114,9 +93,9 @@ int main(int argc, char **argv) {
   Fitness fitness = eval;
   Parameters p = parameter_parse(argc, argv);
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
-  float wi[14] = {0.000000, 0.000100, 0.000000, 0.000100, 0.000000,
-                  0.000100, 0.000000, 1.251300, 0.000000, 0.000100,
-                  0.000000, 0.000100, 0.000000, 0.000100
+  float wi[14] = {0.000000, 2.000000, 0.000000, 0.000100, 0.000000,
+                  0.352596, 2.000000, 2.000000, 0.000000, 0.000100,
+                  0.600357, 0.000000, 0.000000, 0.000100
 
   };
   Weights w = {
@@ -130,8 +109,8 @@ int main(int argc, char **argv) {
       .wl = {wi[10], wi[11]},
       .ll = {wi[12], wi[13]},
   };
-  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 2.0, 0);
-  //float *res = malloc(sizeof(float)*p.population_size); 
+  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 2.0, start_time+rank);
+  // float *res = malloc(sizeof(float)*p.population_size);
   MPI_Barrier(MPI_COMM_WORLD);
 
   if (rank == 0) {

### 2025-08-22 code/bin_mpich_2.c
better parameters

@@ -21,22 +21,23 @@ int main(int argc, char *argv[]) {
 
   // set parameters
   Parameters p = parameter_parse(argc, argv);
-float wi[] = {
-  0.0f,
-  2.0f,
-  0.0f,
-  0.0001f,
-  0.0f,
-  0.352596f,
-  2.0f,
-  2.0f,
-  0.0f,
-  0.0001f,
-  0.600357f,
-  0.0f,
-  0.0f,
-  0.0001f
+float wi[14] = {
+  0.000000,
+2.000000,
+0.000000,
+0.907037,
+0.000000,
+0.000100,
+0.216043,
+0.080545,
+0.000000,
+0.100000,
+0.833333,
+0.336281,
+0.000000,
+0.244496
 };
+
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},

### 2025-08-22 code/bin_training_mpich.c

@@ -93,11 +93,22 @@ int main(int argc, char **argv) {
   Fitness fitness = eval;
   Parameters p = parameter_parse(argc, argv);
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
-  float wi[14] = {0.000000, 2.000000, 0.000000, 0.000100, 0.000000,
-                  0.352596, 2.000000, 2.000000, 0.000000, 0.000100,
-                  0.600357, 0.000000, 0.000000, 0.000100
-
-  };
+  float wi[14] = {
+  0.000000,
+2.000000,
+0.000000,
+0.907037,
+0.000000,
+0.000100,
+0.216043,
+0.080545,
+0.000000,
+0.100000,
+0.833333,
+0.336281,
+0.000000,
+0.244496
+};
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},

### 2025-08-22 code/bin_training_parameters.c

@@ -97,11 +97,22 @@ int main() {
 };
   */
   srand(time(NULL));
-  float best[14] = {0.000000, 0.000100, 0.000000, 0.000100, 0.000000,
-                    0.000100, 0.000000, 1.251300, 0.000000, 0.000100,
-                    0.000000, 0.000100, 0.000000, 0.000100
-
-  };
+  float best[14] = {
+  0.000000,
+2.000000,
+0.000000,
+0.907037,
+0.000000,
+0.000100,
+0.216043,
+0.080545,
+0.000000,
+0.100000,
+0.833333,
+0.336281,
+0.000000,
+0.244496
+};
   unsigned int seed = rand();
   eval(best, &seed, 10);
   // printf("starting: %f\n", best_fitness);

### 2025-08-22 code/bin_mpich.c
cleanup

@@ -21,23 +21,42 @@ int main(int argc, char *argv[]) {
 
   // set parameters
   Parameters p = parameter_parse(argc, argv);
+float wi[14] = {
+  0.000000,
+2.000000,
+0.000000,
+0.907037,
+0.000000,
+0.000100,
+0.216043,
+0.080545,
+0.000000,
+0.100000,
+0.833333,
+0.336281,
+0.000000,
+0.244496
+};
 
   Weights w = {
       // exploring
-      .al = {0.1, 0.1},
-      .cl = {0.7, 0.7},
+      .al = {wi[0], wi[1]},
+      .cl = {wi[2], wi[3]},
       // swarming
-      .sl = {0.1, 0.1},
-      .fl = {1.0, 1.0},
-      .el = {0.0, 0.0},
-      .wl = {0.6, 0.6},
-      .ll = {0.1, 0.1},
+      .sl = {wi[4], wi[5]},
+      .fl = {wi[6], wi[7]},
+      .el = {wi[8], wi[9]},
+      .wl = {wi[10], wi[11]},
+      .ll = {wi[12], wi[13]},
   };
+  ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
+
   int comm_size, rank;
   MPI_Comm_size(MPI_COMM_WORLD, &comm_size);
   MPI_Comm_rank(MPI_COMM_WORLD, &rank);
   Fitness fitness = rastrigin_fitness;
-  float *res = dragonfly_compute(p, w, fitness, comm_size, rank, 100.0, 0);
+
+  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 100.0, start_time);
 
   MPI_Barrier(MPI_COMM_WORLD);
   unsigned int s = 0;

### 2025-08-22 code/bin_mpich_2.c

@@ -1,75 +1,0 @@
-#include <mpi.h>
-#include <stdbool.h>
-#include <stddef.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <time.h>
-#include <unistd.h>
-
-#include "dragonfly-common.h"
-#include "utils.h"
-
-int main(int argc, char *argv[]) {
-  MPI_Init(NULL, NULL);
-  // wait for all the process to start
-  MPI_Barrier(MPI_COMM_WORLD);
-
-  // start clock
-  clock_t start_time;
-  start_time = clock();
-
-  // set parameters
-  Parameters p = parameter_parse(argc, argv);
-float wi[14] = {
-  0.000000,
-2.000000,
-0.000000,
-0.907037,
-0.000000,
-0.000100,
-0.216043,
-0.080545,
-0.000000,
-0.100000,
-0.833333,
-0.336281,
-0.000000,
-0.244496
-};
-
-  Weights w = {
-      // exploring
-      .al = {wi[0], wi[1]},
-      .cl = {wi[2], wi[3]},
-      // swarming
-      .sl = {wi[4], wi[5]},
-      .fl = {wi[6], wi[7]},
-      .el = {wi[8], wi[9]},
-      .wl = {wi[10], wi[11]},
-      .ll = {wi[12], wi[13]},
-  };
-  ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
-
-  int comm_size, rank;
-  MPI_Comm_size(MPI_COMM_WORLD, &comm_size);
-  MPI_Comm_rank(MPI_COMM_WORLD, &rank);
-  Fitness fitness = rastrigin_fitness;
-
-  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 100.0, start_time);
-
-  MPI_Barrier(MPI_COMM_WORLD);
-  unsigned int s = 0;
-  float fit = fitness(res, &s, p.problem_dimensions);
-
-  if (rank == 0) {
-    printf("found fitness=%f\n", fit);
-    for (unsigned int i = 0; i < p.problem_dimensions; i++) {
-      printf("%f\n", res[i]);
-    }
-    double duration = (double)(clock() - start_time) / CLOCKS_PER_SEC;
-    printf("Execution time = %f\n", duration);
-  }
-  free(res);
-  MPI_Finalize();
-}

### 2025-08-22 code/bin_serial.c

@@ -1,67 +1,0 @@
-#include <stdbool.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <time.h>
-
-#include "dragonfly-common.h"
-#include "utils.h"
-
-int main(int argc, char *argv[]) {
-  // start clock
-  clock_t start_time;
-  start_time = clock();
-
-  // set parameters
-  Parameters p = parameter_parse(argc, argv);
-  unsigned int seed = time(NULL);
-  float *shifted_tmp = malloc(sizeof(float) * p.problem_dimensions);
-  float *shifted_rotation =
-      malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
-  float *shifted_shift = init_array(p.problem_dimensions, 100.0, &seed);
-  init_matrix(shifted_rotation, 100.0, p.problem_dimensions, &seed);
-
-  init_shifted_fitness(shifted_tmp, shifted_rotation, shifted_shift,
-                       rastrigin_fitness);
-
-  Fitness fitness = shifted_fitness;
-  float wi[14] = {0.803008, 0.801540, 0.000000, 4.131542, 1.180980,
-                  2.213441, 1.289022, 0.828981, 0.100000, 0.022841,
-                  0.000000, 0.000000, 0.108948, 0.404676};
-  Weights w = {
-      // exploring
-      .al = {wi[0], wi[1]},
-      .cl = {wi[2], wi[3]},
-      // swarming
-      //.sl = {wi[4], wi[5]},
-      .fl = {wi[6], wi[7]},
-      .el = {wi[8], wi[9]},
-      .wl = {wi[10], wi[11]},
-      .ll = {wi[12], wi[13]},
-  };
-
-  float *res = dragonfly_compute(p, w, fitness, 1, 0, 100.0, seed);
-  unsigned int s = 0;
-  float fit = fitness(res, &s, p.problem_dimensions);
-
-  printf("found fitness=%f\n", fit);
-  for (unsigned int i = 0; i < p.problem_dimensions; i++) {
-    printf("%f\n", res[i]);
-  }
-  free(res);
-  free(shifted_rotation);
-  free(shifted_shift);
-  free(shifted_tmp);
-  double duration = (double)(clock() - start_time) / CLOCKS_PER_SEC;
-  printf("Execution time = %f\n", duration);
-}
-
-unsigned int raw_sendrecv_shift;
-void raw_sendrecv(Message *send, unsigned int destination, Message *recv_buffer,
[…]

### 2025-08-22 code/dragonfly-common.c

@@ -233,9 +233,6 @@ unsigned int assing_logical_chunks(unsigned int phisical_chunk_start,
         .to_shift_right = !starts_in_previous && ends_in_next,
         .complete = !starts_in_previous && !ends_in_next,
     };
-    // printf("chunk: %d-%d %d-%d %d %d %d\n", phisical_chunk_start,
-    //          phisical_chunk_end, cur_start, cur_end, cur.to_shift_left,
-    //          cur.to_shift_right, index);
     buffer[index] = cur;
     index++;
     cur_start += logical_chunk_size;
@@ -268,15 +265,11 @@ void comunicate(LogicalChunk *cur, MPI_Datatype type,
 
   ComputationStatus tmp;
   unsigned inner_rank_id = rank_id - start;
-  // printf("comunicate1 %d %d %d %d %d\n", cur->start, cur->end, size, rank_id,
-  // required_steps);
   for (unsigned int i = 0; i < required_steps; i++) {
 
     unsigned partner = (inner_rank_id ^ (1 << i)) + start;
 
     if (start <= partner && partner < end) {
-      // printf("comunicate2 %d %d %d->%d\n", start, end, rank_id, partner);
-      // fflush(stdout);
       MPI_Sendrecv(&cur->comp, 1, type, partner, 1, &tmp, 1, type, partner, 1,
                    MPI_COMM_WORLD, MPI_STATUS_IGNORE);
       computation_status_merge(&cur->comp, &tmp, dim);
@@ -290,20 +283,14 @@ void comunicate(LogicalChunk *cur, MPI_Datatype type,
   for (int i = required_steps - 2; i >= 0; i--) {
     unsigned partner = (inner_rank_id ^ (1 << i)) + start;
     if (start <= partner && partner < end) {
-      // printf("comunicate3 %d %d %d->%d\n", start, end, rank_id, partner);
-      // fflush(stdout);
       MPI_Sendrecv(&cur->comp, 1, type, partner, 1, &tmp, 1, type, partner, 1,
                    MPI_COMM_WORLD, MPI_STATUS_IGNORE);
 
-      // computation_status_merge(&cur->comp, &tmp, dim);
       if (cur->comp.n < tmp.n) {
         memcpy(&cur->comp, &tmp, sizeof(ComputationStatus));
       }
     }
   }
-
-  // printf("end comunicate1 %d %d %d %d %d\n", cur->start, cur->end, size,
-  // rank_id, required_steps);
 }
 
 // TODO paralelize
[…]

### 2025-08-22 code/utils.c

@@ -170,410 +170,6 @@ float shifted_fitness(float *inp, unsigned int *seed, unsigned int dim) {
 #include <omp.h>
 
 #endif
-/*
-void inner_dragonfly_step(Dragonfly *d, float *average_speed,
-                          float *cumulated_pos, float *food, float *enemy,
-                          unsigned int N, unsigned int base, unsigned int limit,
-                          unsigned int random_seed) {
-  unsigned int dimensions = d->dim;
-  float S;
-  float A;
-  float C;
-  float F;
-  float E;
-  float levy;
-
-  float *cur_pos;
-  float *cur_speed;
-
-  for (unsigned int j = base; j < limit; j++) {
-    unsigned random = random_seed + j;
-    cur_pos = d->positions + j * dimensions;
-    cur_speed = d->speeds + j * dimensions;
-
-    // compute speed = sSi + aAi + cCi + fFi + eEi + w
-
-   for (unsigned int i = 0; i < dimensions; i++) {
-      S = (cumulated_pos[i] / ((float)N)) - cur_pos[i];
-      A = average_speed[i];
-      C = (cumulated_pos[i] / (float)N) - cur_pos[i];
-      F = food[i] - cur_pos[i];
-      E = enemy[i] + cur_pos[i];
-      levy = RAND_FLOAT(1.0, &random);
-
-      cur_speed[i] *= d->w.w;
-      cur_speed[i] += d->w.s * S;
-      cur_speed[i] += d->w.a * A;
-      cur_speed[i] += d->w.c * C;
-      cur_speed[i] += d->w.f * F;
-      cur_speed[i] += d->w.e * E;
-      cur_speed[i] += levy;
-
-      cur_pos[i] += cur_speed[i];
-    }
-  }
-}
-
-void dragonfly_compute_step(Dragonfly *d, float *average_speed,
-                            float *cumulated_pos, float *food, float *enemy,
-                            unsigned int N, unsigned int NR_THREADS) {
-  unsigned int base_random = rand_r(&d->seed);
-  //printf("base_random=%d\n", base_random);
-  // if more than 1 thread
-  unsigned int rest = d->local_n % NR_THREADS;
-  unsigned int ratio = d->local_n / NR_THREADS;
-  if (NR_THREADS == 1) {
-    inner_dragonfly_step(d, average_speed, cumulated_pos, food, enemy, N, 0,
-                         d->local_n, base_random);
-  } else {
-    
[…]

### 2025-08-22 code/bin_mpich.c
better parameters

@@ -23,19 +23,19 @@ int main(int argc, char *argv[]) {
   Parameters p = parameter_parse(argc, argv);
 float wi[14] = {
   0.000000,
-2.000000,
-0.000000,
-0.907037,
+0.000100,
+1.023072,
+1.628618,
 0.000000,
 0.000100,
-0.216043,
-0.080545,
+0.930669,
+1.020353,
 0.000000,
 0.100000,
-0.833333,
-0.336281,
 0.000000,
-0.244496
+0.000100,
+0.000000,
+0.000100
 };
 
   Weights w = {

### 2025-08-22 code/bin_training_mpich.c

@@ -14,8 +14,8 @@ float eval(float *wi, unsigned int *seedi, unsigned int d) {
     if (wi[i] < 0.0) {
       wi[i] = 0.0;
     }
-    if (wi[i] > 2.0) {
-      wi[i] = 2.0;
+    if (wi[i] > 4.0) {
+      wi[i] = 4.0;
     }
     if (i % 2 == 0 && wi[i] == wi[i + 1]) {
       wi[i + 1] += 0.0001;
@@ -95,19 +95,19 @@ int main(int argc, char **argv) {
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
   float wi[14] = {
   0.000000,
-2.000000,
-0.000000,
-0.907037,
+0.000100,
+1.023072,
+1.628618,
 0.000000,
 0.000100,
-0.216043,
-0.080545,
+0.930669,
+1.020353,
 0.000000,
 0.100000,
-0.833333,
-0.336281,
 0.000000,
-0.244496
+0.000100,
+0.000000,
+0.000100
 };
   Weights w = {
       // exploring

### 2025-08-23 code/dragonfly-common.c
implemented first parallelization version with openmp

@@ -18,7 +18,7 @@ Parameters parameter_parse(int argc, char *argv[]) {
   if (argc != 6) {
     fprintf(stderr, "invalid parameter count: expected population_size, "
                     "starting_chunk_count, iterations, problem_dimensions, "
-                    "threads_per_process");
+                    "threads_per_process\n");
     exit(-1);
   }
   p.population_size = atoi(argv[1]);
@@ -196,15 +196,7 @@ void Build_mpi_type_computation_status(MPI_Datatype *data) {
 
 // #endif
 
-typedef struct {
-  unsigned int start;
-  unsigned int end;
-  bool to_shift_left;
-  bool to_shift_right;
-  bool complete;
 
-  ComputationStatus comp;
-} LogicalChunk;
 
 unsigned int assing_logical_chunks(unsigned int phisical_chunk_start,
                                    unsigned int phisical_chunk_end,
@@ -293,56 +285,7 @@ void comunicate(LogicalChunk *cur, MPI_Datatype type,
   }
 }
 
-// TODO paralelize
-// TODO move to right place
-//  it computes the best, the food, the enemy, and the sums of speeds and
-//  positions of an interval. the interval must be inside the current thread
-//  chunk
-void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
-                                unsigned int *seed) {
 
-  unsigned start = max(current_chunk->start, d->start);
-  unsigned end = min(current_chunk->end, d->end);
-
-  unsigned dim = d->dim;
-  float *cumulated_pos = current_chunk->comp.cumulated_pos;
-  float *cumulated_speed = current_chunk->comp.cumulated_speeds;
-
-  zeroed(cumulated_pos, dim);
-  zeroed(cumulated_speed, dim);
-
-  float next_enemy_fitness = d->fitness(d->positions, &d->seed, dim);
-  float next_food_fitness = next_enemy_fitness;
-  // status->n = d->local_n;
-
-  unsigned int indexes[2] = {0, 0};
-  end = end - d->start;
-  for (unsigned int k = start - d->start; k < end; k++) {
-    float *iter_pos = d->positions + dim * k;
-    float *iter_speed = d->speeds + dim * k;
-    sum_assign(cumulated_pos, iter_pos, dim);
-    sum_assign(cumulated_speed, iter_speed, dim);
-
-    float fitness = d->fitness(iter_pos, seed, dim);
-
[…]

### 2025-08-23 code/dragonfly-common.h

@@ -1,6 +1,7 @@
 #ifndef DA_COMMON
 #define DA_COMMON
 #include <math.h>
+#include <stdbool.h>
 #define MESSAGE_SIZE 16
 #define MAX_CHUNKS 4048
 
@@ -80,6 +81,16 @@ typedef struct {
 
   unsigned int n;
 } ComputationStatus;
+
+typedef struct {
+  unsigned int start;
+  unsigned int end;
+  bool to_shift_left;
+  bool to_shift_right;
+  bool complete;
+
+  ComputationStatus comp;
+} LogicalChunk;
 
 typedef struct {
   unsigned int population_size, starting_chunk_count, iterations,

### 2025-08-23 code/utils.c

@@ -40,24 +40,36 @@ void zeroed(float *dest, unsigned int size) {
 }
 
 void sum_assign(float *dest, float *source, unsigned int size) {
+#ifdef DUSE_OPENMP
+#pragma omp parallel for reduction(+: dest)
+#endif
   for (unsigned int i = 0; i < size; i++) {
     dest[i] += source[i];
   }
 }
 
 void sub_assign(float *dest, float *source, unsigned int size) {
+#ifdef DUSE_OPENMP
+#pragma omp parallel for reduction(-: dest)
+#endif
   for (unsigned int i = 0; i < size; i++) {
     dest[i] -= source[i];
   }
 }
 
 void scalar_prod_assign(float *dest, float val, unsigned int size) {
+#ifdef DUSE_OPENMP
+#pragma omp parallel for reduction(*: dest)
+#endif
   for (unsigned int i = 0; i < size; i++) {
     dest[i] *= val;
   }
 }
 
 float dot_product(float *a, float *b, unsigned int size) {
+#ifdef DUSE_OPENMP
+#pragma omp parallel for reduction(+: ret)
+#endif
   float ret = 0.0;
   for (unsigned int i = 0; i < size; i++) {
     ret += a[i] * b[i];
@@ -67,7 +79,6 @@ float dot_product(float *a, float *b, unsigned int size) {
 
 float length(float *inp, unsigned int size) {
   float ret = dot_product(inp, inp, size);
-
   return sqrt(ret);
 }
 
@@ -166,10 +177,158 @@ float shifted_fitness(float *inp, unsigned int *seed, unsigned int dim) {
   return FITNESS(TMP, seed, dim);
 }
 
+void inner_computation_accumulate(Dragonfly* d,unsigned int thread_start,unsigned int thread_end,
+																	float* local_cumulated_pos,float* local_cumulated_speed,
+																	float* local_food_fitness,float* local_enemy_fitness,
+																	unsigned int* local_indexes,unsigned int dim,unsigned int *local_seed){
+	for (unsigned int k = thread_start; k < thread_end; k++) {
+		float *iter_pos = d->positions + dim * k;
+		float *iter_speed = d->speeds + dim * k;
+		sum_assign(local_cumulated_pos, iter_pos, dim);
+		sum_assign(local_cumulated_speed, iter_speed, dim);
+
+		float fitness = d->fitness(iter_pos,local_seed, dim);
+
+		if (fitness > *local_food_fitness) {
[…]

### 2025-08-23 code/utils.h

@@ -8,6 +8,10 @@
 #define min(a, b) ((a) < (b) ? (a) : (b))
 #define max(a, b) ((a) > (b) ? (a) : (b))
 
+#ifdef USE_OPENMP
+#include <omp.h>
+#endif
+
 float *init_array(unsigned int dimensions, float range_max, unsigned int *seed);
 void sum_assign(float *dest, float *source, unsigned int size);
 void sub_assign(float *dest, float *source, unsigned int size);
@@ -29,6 +33,9 @@ void init_matrix(float *inp, float range_max, unsigned int dim, unsigned int *se
 void dragonfly_compute_step(Dragonfly *d, float *average_speed,
     float *cumulated_pos, float *food, float *enemy,
     unsigned int N, unsigned int NUM_THREADS);
-void computation_accumulate(ComputationStatus *message, Dragonfly *d, float* best, float* best_fitness, unsigned int NUM_THREADS);
+void new_computation_accumulate(Dragonfly *d, 
+																LogicalChunk *current_chunk,
+                                unsigned int *seed,
+																unsigned int nr_threads);
 
 #endif

### 2025-08-24 text-paper/paper.tex
introduction

@@
   \section*{Abstract}
-  \textbf{Dragonfly algorithm is a new kind of approach used to solve optimization problems through the emulation of dragonflies movements, but it lacks scalability.
-In this paper the authors try to set up an approach based on parallelization thanks to the use of API for high performance computing like MPI and OpenMP, then there will be a comparison of execution times between presented implementation and the classic approach based on serial programming, in order to see if parallelization could lead to a performance improvement and possible future extensions.}\newline]
+  \textbf{
+    Dragonfly algorithm is an approach used to solve optimization problems through the emulation of dragonflies movements, but it lacks scalability.
+  In this paper the authors propose an approach based on parallelization using dedicated libraries for high performance multiprocess applications like MPI, and OpenMP API for multithread parallelization.
+  A comparison of execution times is then conducted between presented implementation and the classic approach based on serial programming,
+  to determine whether parallelization leads to performance improvements and enables possible future extensions.
+  }\newline]
 
 \section{Introduction}
+Swarm intelligence algorithms are inspired by the collective behavior of biological systems, where simple rules followed by individuals lead to the emergence of complex and adaptive group behavior.
+The concept was first introduced in Reynolds' Boids model (1987) \cite{TODO}, which simulated flocking behavior, and gained significant traction in Computer Science with the Ant Colony Optimization algorithm (1995) \cite{TODO}.
+These algorithms are meta-heuristic in nature, leveraging decentralized and parallel exploration-exploitation strategies to solve optimization problems.
 
-Recent years have seen the raise and evolution of swarm intelligent algorithms, based on meta-heuristic approach where single individuals of a complex population follow an exploration-exploitation methodology to find the best solution in their local search space;
-each member of the group work independently from the others, leading to a subdivision of the labour which decrease the overall amount of execution time, working as decentralized and parallel units; but also, due to the fact that they mimic the behaviour of biological life forms, it is possible then that they could generate random patterns through the interaction with each other.
-The dragonfly algorithm has been proposed for the first time in 2016, and its purpose is to simulate the behaviour of a dragonfly population according to the principles of swarm intelligent algorihms seen previously.
-It has gained more and more popularity: it is in estimated that during 2019 more than 300 scientific papers cited the algorithm \cite{DASurvey}, due to the fact that it has demonstrated some characteristics similar to other algoritms of the same kind, however in contrast with other swarm intelligent algorithms however it present less probability to be trapped into a local optima \cite{DAReview}.
-This paper aims to define a new approach based on popular parallel APIs, and also trying to understand if parallelization could lead to a significant speed up in performances that can be considered by future researches on this algorithm
+The dragonfly algorithm, first proposed in 2014 and published in 2015 \cite{DAReview}, is a bio-inspired optimization algorithm that simulates the behavior of dragonfly swarms.
+By mimicking natural behaviors such as alignment, separation, cohesion, attraction to food sources, and distraction from enemies, the algorithm achieves a balance between exploration and exploitation.
+This makes it particularly effective at navigating complex search spaces while reducing the likelihood of being trapped in local optima.
+As of 2019, the algorithm has been cited in over 300 scientific papers, highlighting its impact and popularity in the field.
+
+This paper focuses on parallelizing the dragonfly algorithm to improve its scalability and performance.
+Using popular parallel APIs such as OpenMP for multithreading and MPI for distributed memory systems, we aim to demonstrate how parallelization can significantly reduce execution time while maintaining the algorithm's effectiveness.
+The results of this study could pave the way for future research and applications of the dragonfly algorithm in high-performance computing environments.
 
 \subsection*{How the algorithm works}
-The execution begins with the initialization of some arrays which are the position $P$ of all dragonfiles and the `step` vector $\Delta P$ representing the velocities of each single individual,the position of source foods $F$ and the position of enemies $E$;
-then the algorithm executes an iteration loop where, for each dragonfly, we calculate some coefficient that will be used to update the internal state of the program.
-This procedure is simulated through the observation of repetitive patterns observed in the behaviour of dragonflies, which are:
+The execution of the dragonfly algorithm begins with the initialization of several arrays: the position vector $P$ representing the positions of all dragonflies, the step vector $\Delta P$ representing their velocities, the position of food sources $F$, and the position of enemies $E$.
+The algorithm then enters an iterative loop where, for each dragonfly, coefficients are calculated to update its internal state.
+These coefficients are derived from the observed behaviors of dragonflies in nature, which include:
 
 \begin{itemize}
-  \item Alignment of their velocity with other members of the same group, given by the formula  $A_i = \frac{\sum^M_{j=1}V_j}{M}$, where $V_j$ is the velocity of the j-th neighbour
+  \item \textbf{Alignment}: The tendency to match the velocity of neighboring dragonflies, calculated as:
+  $$A_i = \frac{\sum_{j=1}^M V_j}{M},$$
+  where $V_j$ is the velocity of the $j$-th neighbor and $M$ is the number of neighbors.
 
-  \item Separation of single members in order to avoid collision, and it is retrieved by the equation  $S_i = - \sum^M_{j=1}{P - P_j}$, where $P_j$ is the position of the j-th neighbour and $P$ the current individual position.
+  \item \textbf{Separation}: The tendency to avoid collisions with other dragonflies, given by:
+  $$S_i = -\sum_{j=1}^M (P - P_j),$$
+  where $P_j$ is the position of the $j$-th neighbor and $P$ is the current dragonfly's position.
 
-  \item Cohesion towards the center of the group, given by $C_i = \frac{\sum^M_{j=1}P_j}{M} - P$ 
-  \item Attraction to a common source food, expressed by the formula $ F_i = F^+ - P$, where $F^+$ is the position of the current prey and $P$ is the position of the current dragonfly
-  \item Distraction of enemies, represented by the equation $ E_i = E^- + P$ where $E^-$ is the position of the current enemy 
+  \item \textbf{Cohesion}: The tendency to move toward the center of the group, expressed as:
+  $$C_i = \frac{\sum_{j=1}^M P_j}{M} - P.$$
+
+  \item \textbf{Attraction to food}: The tendency to move toward a food source, represented by:
+  $$F_i = F^+ - P,$$
+  where $F^+$ is the position of the food source.
+
+  \item \textbf{Distraction from enemies}: The tendency to move away from enemies, modeled as:
+  $$E_i = E^- + P,$$
+  where $E^-$ is the position of the enemy.
 \end{itemize}
 
-\noindent Once all the parameters have been calculated we can now update the current step vector of each dragonfly using the following equation: $$\Delta P^{t+1}_{i} = (sS_i + aA_i + fF_i + cC_i) + \omega\Delta P^t_i$$ where $s,a,f,c$ are respectevly the separation weight, the alignment weight, the attraction weight, the cohesion weight and $\omega$ is the inertia weight, all of them are randomly chosen in the interval $[0,1]$.
-When all step vectors have been updated we can now set the new position of the dragonflies using this equation, which takes the new step vector as an input: $$P^{t+1}_{i} = P_i^t + \Delta P_{i}^{t+1}$$However if there is a dragonfly which does not have any neighbour then we assume that the overall population moves in a random way: the most common method to do that is to use a random walk function.
-The previous formula then must be modified into this one here: $$P^{t+1}_i = P_i^t + Levy(d)$$ where $d$ is the dimension of the position vector, and $Levy(d)$ is the Levy flight function, which is calculated in this way: $$Levy(d) = 0.01\times\frac{r_1 \times \sigma}{|r_2|^{\frac{1}{\beta}}}$$ where $r1$ and $r2$ are two random numbers uniformally distributed in the interval $[0,1]$, and $\sigma$ is a vector calculated using the following formula: $$\sigma = \frac{sin(\frac{\beta\pi}{2})\times\Gamma(1+\beta)}{\Gamma(\frac{\beta+1}{2})\times \beta \times 2^{(\beta-1)/2}}$$ where $\Gamma(x)$ is the gamma function \cite{WikiGamma}, and $\beta$ is a constant.Another valid approach is to apply, instead of Levy flight, the brownian motion model, used especially in gas and particle analysis: this seems to return better performances \cite{BDragonfly}.
+\noindent Once these coefficients are calculated, the step vector $\Delta P$ for each dragonfly is updated using the formula:
+$$\Delta P_i^{t+1} = sS_i + aA_i + fF_i + cC_i + \omega \Delta P_i^t,$$
+where $s$, $a$, $f$, $c$, and $\omega$ are weights for separation, alignment, attraction, cohesion, and inertia, respectively.
+These weights are randomly chosen within the interval $[0,1]$.
+The new position of each dragonfly is then computed as:
+$$P_i^{t+1} = P_i^t + \Delta P_i^{t+1}.$$
+
+If a dragonfly has no neighbors, it is assumed to move randomly. This is modeled using a random walk function, often implemented as a Lévy flight:
+$$P_i^{t+1} = P_i^t + \mathrm{Levy}(d),$$
+where $d$ is the dimension of the position vector. The Lévy flight function is defined as:
+$$\mathrm{Levy}(d) = 0.01 \times \frac{r_1 \times \sigma}{|r_2|^{1/\beta}},$$
+where $r_1$ and $r_2$ are random numbers uniformly distributed in $[0,1]$, and $\sigma$ is calculated as:
+$$\sigma = \frac{\sin\left(\frac{\beta \pi}{2}\right) \Gamma(1+\beta)}{\Gamma\left(\frac{\beta+1}{2}\right) \beta 2^{(\beta-1)/2}},$$
+with $\Gamma(x)$ being the gamma function and $\beta$ a constant. Alternatively, Brownian motion can be used instead of Lévy flight, which has shown better performance in some cases \cite{BDragonfly}.
 

### 2025-08-24 code/bin_training_mpich.c

@@ -93,22 +93,9 @@ int main(int argc, char **argv) {
   Fitness fitness = eval;
   Parameters p = parameter_parse(argc, argv);
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
-  float wi[14] = {
-  0.000000,
-0.000100,
-1.023072,
-1.628618,
-0.000000,
-0.000100,
-0.930669,
-1.020353,
-0.000000,
-0.100000,
-0.000000,
-0.000100,
-0.000000,
-0.000100
-};
+  float wi[14] = {0.000000, 0.000100, 1.023072, 1.628618, 0.000000,
+                  0.000100, 0.930669, 1.020353, 0.000000, 0.100000,
+                  0.000000, 0.000100, 0.000000, 0.000100};
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},
@@ -120,7 +107,8 @@ int main(int argc, char **argv) {
       .wl = {wi[10], wi[11]},
       .ll = {wi[12], wi[13]},
   };
-  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 2.0, start_time+rank);
+  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 2.0,
+                                 start_time + rank);
   // float *res = malloc(sizeof(float)*p.population_size);
   MPI_Barrier(MPI_COMM_WORLD);
 

### 2025-08-24 code/dragonfly-common.c
fixed segfault problem

@@ -7,6 +7,8 @@
 #include <stdbool.h>
 #include <stddef.h>
 #include <stdio.h>
+
+#include <time.h>
 // #ifdef USE_MPI
 #include <mpi.h>
 #include <unistd.h>
@@ -417,7 +419,7 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
     //  1) accumulate
     for (unsigned int j = 0; j < n_local_chunks; j++) {
       // TODO paralelization opportunity if handling correctly random seed
-      new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,threads);
+      new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,p.threads_per_process);
     }
     if (threads > 1) {
 
@@ -556,3 +558,206 @@ void update_chunk_size(ChunkSize *c) {
                (float)c->start_count;
   c->count = (unsigned int)(round(temp));
 }
+
+void inner_computation_accumulate(Dragonfly* d,unsigned int thread_start,unsigned int thread_end,
+																	float* local_cumulated_pos,float* local_cumulated_speed,
+																	float* local_food_fitness,float* local_enemy_fitness,
+																	unsigned int* local_indexes,unsigned int dim,unsigned int *local_seed){
+	for (unsigned int k = thread_start; k < thread_end; k++) {
+		float *iter_pos = d->positions + dim * k;
+		float *iter_speed = d->speeds + dim * k;
+		sum_assign(local_cumulated_pos, iter_pos, dim);
+		sum_assign(local_cumulated_speed, iter_speed, dim);
+
+		float fitness = d->fitness(iter_pos,local_seed, dim);
+
+		if (fitness > *local_food_fitness) {
+			local_indexes[0] = k;
+			*local_food_fitness = fitness;
+		}
+		if (fitness < *local_enemy_fitness) {
+			local_indexes[1] = k;
+			*local_enemy_fitness = fitness;
+		}
+	}
+}
+
+void update_status(Dragonfly* d,LogicalChunk* current_chunk,
+									 float next_food_fitness,float next_enemy_fitness,
+									 unsigned int* indexes,unsigned int start,
+									 unsigned int end,unsigned int dim){
+
+	memcpy(current_chunk->comp.next_food, d->positions + indexes[0] * dim,
+         sizeof(float) * dim);
[…]

### 2025-08-24 code/utils.c

@@ -40,40 +40,29 @@ void zeroed(float *dest, unsigned int size) {
 }
 
 void sum_assign(float *dest, float *source, unsigned int size) {
-#ifdef DUSE_OPENMP
-#pragma omp parallel for reduction(+: dest)
-#endif
   for (unsigned int i = 0; i < size; i++) {
     dest[i] += source[i];
   }
 }
 
 void sub_assign(float *dest, float *source, unsigned int size) {
-#ifdef DUSE_OPENMP
-#pragma omp parallel for reduction(-: dest)
-#endif
   for (unsigned int i = 0; i < size; i++) {
     dest[i] -= source[i];
   }
 }
 
 void scalar_prod_assign(float *dest, float val, unsigned int size) {
-#ifdef DUSE_OPENMP
-#pragma omp parallel for reduction(*: dest)
-#endif
   for (unsigned int i = 0; i < size; i++) {
     dest[i] *= val;
   }
 }
 
 float dot_product(float *a, float *b, unsigned int size) {
-#ifdef DUSE_OPENMP
-#pragma omp parallel for reduction(+: ret)
-#endif
   float ret = 0.0;
   for (unsigned int i = 0; i < size; i++) {
     ret += a[i] * b[i];
   }
+
   return ret;
 }
 
@@ -126,7 +115,6 @@ float levy() {
 }*/
 
 void brownian_motion(float *inp, unsigned int dim, unsigned int *seed) {
-
   for (unsigned int i = 0; i < dim; i++) {
     inp[i] = RAND_FLOAT(1.0, seed);
   }
@@ -177,158 +165,5 @@ float shifted_fitness(float *inp, unsigned int *seed, unsigned int dim) {
   return FITNESS(TMP, seed, dim);
 }
 
-void inner_computation_accumulate(Dragonfly* d,unsigned int thread_start,unsigned int thread_end,
-																	float* local_cumulated_pos,float* local_cumulated_speed,
-																	float* local_food_fitness,float* local_enemy_fitness,
-																	unsigned int* local_indexes,unsigned int dim,unsigned int *local_seed){
-	for (unsigned int k = thread_start; k < thread_end; k++) {
-		float *iter_pos = d->positions + dim * k;
-		float *iter_speed = d->speeds + dim * k;
-		sum_assign(local_cumulated_pos, iter_pos, dim);
-		sum_assign(local_cumulated_speed, iter_speed, dim);
-
-		float fitness = d->fitness(iter_pos,local_seed, dim);
-
[…]

### 2025-08-25 code/dragonfly-common.c
removed local buffers

@@ -640,63 +640,43 @@ void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
 								end,dim);
 		return ; 
 	}
-#pragma omp parallel num_threads(nr_threads) schedule(runtime) 
+#pragma omp parallel num_threads(nr_threads)
 	{
 		unsigned int thread_id = omp_get_thread_num();
 		unsigned int thread_start = max(start,start+(thread_id*elems_per_thread));
 		unsigned int thread_end = min(thread_start + elems_per_thread,end);
 
-	  unsigned int local_indexes[2]; 
-		float local_fitnesses[2];
-		float local_cumulated_pos[dim];
-		float local_cumulated_speed[dim];
+		unsigned int local_seed = (*seed) * (thread_id+1);
 
-		local_fitnesses[0] = next_food_fitness;
-		local_fitnesses[1] = next_enemy_fitness;
-		unsigned int local_seed = (*seed) + thread_id;
+		zeroed(thread_cumulated_pos[thread_id],dim);
+		zeroed(thread_cumulated_speed[thread_id],dim);
 
-		local_indexes[0] = thread_start;
-		local_indexes[1] = thread_start;
+		thread_indexes[thread_id][0] = thread_start;
+		thread_indexes[thread_id][1] = thread_start;
 
-		zeroed(local_cumulated_pos,dim);
-		zeroed(local_cumulated_speed,dim);
-
-		float positions[elems_per_thread][dim];
-		float speeds[elems_per_thread][dim];
-
-		for(unsigned int i=0; i<elems_per_thread; i++){
-			for(unsigned int j=0; j<dim; j++){
-				memcpy(positions[i],d->positions + i*dim + j,dim*sizeof(float));
-				memcpy(speeds[i],d->speeds + i*dim + j,dim*sizeof(float));
-			}
-		}
+		thread_fitnesses[thread_id][0] = next_food_fitness;
+		thread_fitnesses[thread_id][1] = next_enemy_fitness;
 
 		for (unsigned int k = thread_start; k < thread_end; k++) {
-			float *iter_pos = positions[k];
-			float *iter_speed = speeds[k];
+			float *iter_pos = d->positions + dim*k;
+			float *iter_speed = d->speeds + dim*k;
 
 			for(unsigned int i=0; i<dim; i++){
-				local_cumulated_pos[i] += iter_pos[i];
[…]

### 2025-08-26 code/dragonfly-common.c
added local buffers to decrease global memory accesses

@@ -648,35 +648,40 @@ void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
 
 		unsigned int local_seed = (*seed) * (thread_id+1);
 
-		zeroed(thread_cumulated_pos[thread_id],dim);
-		zeroed(thread_cumulated_speed[thread_id],dim);
+		unsigned int local_indexes[2] = {thread_start,thread_start};
+		float local_fitnesses[2] = {next_food_fitness,next_enemy_fitness};
+		float local_cumulated_pos[dim];
+		float local_cumulated_speed[dim];
 
-		thread_indexes[thread_id][0] = thread_start;
-		thread_indexes[thread_id][1] = thread_start;
-
-		thread_fitnesses[thread_id][0] = next_food_fitness;
-		thread_fitnesses[thread_id][1] = next_enemy_fitness;
-
+		zeroed(local_cumulated_pos,dim);
+		zeroed(local_cumulated_speed,dim);
+#pragma omp for
 		for (unsigned int k = thread_start; k < thread_end; k++) {
 			float *iter_pos = d->positions + dim*k;
 			float *iter_speed = d->speeds + dim*k;
 
 			for(unsigned int i=0; i<dim; i++){
-				thread_cumulated_pos[thread_id][i] += iter_pos[i];
-				thread_cumulated_speed[thread_id][i] += iter_speed[i];
+				local_cumulated_pos[i] += iter_pos[i];
+				local_cumulated_speed[i] += iter_speed[i];
 			}
 
 			float fitness = d->fitness(iter_pos,&local_seed, dim);
 
-			if (fitness > thread_fitnesses[thread_id][0]) {
-				thread_indexes[thread_id][0] = k;
-				thread_fitnesses[thread_id][0] = fitness;
+			if (fitness > local_fitnesses[0]) {
+				local_indexes[0] = k;
+				local_fitnesses[0] = fitness;
 			}
-			if (fitness < thread_fitnesses[thread_id][1]) {
-				thread_indexes[thread_id][1] = k;
-				thread_fitnesses[thread_id][1] = fitness;
+			if (fitness < local_fitnesses[1]) {
+				local_indexes[1] = k;
+				local_fitnesses[1] = fitness;
 			}
 		}
+
+		memcpy(thread_cumulated_pos[thread_id],local_cumulated_pos,dim*sizeof(float));
[…]

### 2025-08-27 text-paper/paper.tex
mpi paper improvements

@@
 \section{General approach}
-
-Now that we have seen the overall procedure it is possible to write a naive implementation of the algorithm in order to 
-have a base for further developments. An example of the implementation can be observed in the following code snippet:
+With the overall procedure outlined, we can now present a naive implementation of the dragonfly algorithm.
+This initial implementation serves as a foundational baseline for further optimization and parallelization.
+The following code snippet illustrates the core structure of the algorithm:
 
@@
 
-\noindent After the initialization of velocities and positions of agents,
-which will be represented by two separated matricies with the same dimension equal to the number of members of our
-population times the dimension of our problem, the algorithm starts to execute a for loop 
-with a fixed amount of iteration stored in the variable $d\rightarrow N$,  retrieve
-the position and the velocity of the current dragonfly, represented by pointers
-\textit{cur\_pos} and \textit{cur\_speed},
-and the cumulated coordinates of its neighbours. 
-The execution then proceed with the following steps:
+\noindent After initializing the velocities and positions of agents,
+represented by two separate matrices with dimensions equal to the number of agents in the population times the problem's dimensionality, the algorithm begins its main execution loop.
+This loop runs for a fixed number of iterations, stored in the variable $d\rightarrow N$.
+During each iteration, the algorithm retrieves the position and velocity of the current dragonfly, represented by the pointers \textit{cur\_pos} and \textit{cur\_speed}, along with the cumulated coordinates of its neighbors.
+The execution proceeds with the following steps:
 
 \begin{itemize}
-  \item retrieve current speed and position of the j-th dragonfly
-  \item compute separation, alignment, cohesion, food attraction and enemy distraction vectors,
-  stored into local variables named with their abrreviations
-  \item calculate product between weigths and and the single coordinates of the velocity array
-  \item update speed and position the dragonfly considering also brownian motion
+  \item Retrieve the current speed and position of the $j$-th dragonfly.
+  \item Compute the separation, alignment, cohesion, food attraction, and enemy distraction vectors, storing them in local variables with their respective abbreviations.
+  \item Calculate the product of the weights and the individual coordinates of the velocity array.
+  \item Update the speed and position of the dragonfly, incorporating Brownian motion.
 \end{itemize}
 
-\noindent Notice how the computation
-of velocities and speed coordinates is compacted into the same \textbf{for} loop;
-this allows to reduce time execution, because the overall calculation is made locally
-all in the same place, leaving out the necessity of using auxiliary arrays for the coefficients.
-The reason why we divide the separation coefficient for the number of neighbours 
-will be explained later in this paper
+\noindent Notice that the computation of velocities and position updates is compacted into the same \textbf{for} loop.
+This optimization reduces execution time by performing all calculations locally within the loop, eliminating the need for auxiliary arrays to store intermediate coefficients.
+The rationale for dividing the separation coefficient by the number of neighbors will be explained later in this paper (TODO reference).
 
-\subsection*{chunks}
+\section{Critical analysis of the coefficients}
+During our implementation we discovered some ambiguities in the original paper:
+\begin{itemize}
+  \item In the original paper \cite{Original}, the term "neighbors" is used frequently without providing a precise definition.
+  While a general distance metric, such as Euclidean distance, could be assumed, the paper also claims that the algorithm runs in linear time with respect to the number of dragonflies and the iteration count.
+  This claim would not hold if Euclidean distance were used, as calculating the distance between each pair of dragonflies would result in a time complexity of $O(N^2)$.
 
-As anticipated earlier dragonfly algorithm, as many other swarm optimization solutions,
-has a good chance to be stuck into a local minima, without the possibility to find better
-candidates. We can avoid this and improve this approach introducing groupments of dragonflies into chunks, 
-calculating local minima of different areas and compare them in order
-to get as close as possible to the global solution.
+  A possible solution to achieve linear time complexity might involve using advanced geospatial algorithms, but we doubt this was the original intention of the authors.
+  To address this ambiguity, we decided to divide the population into chunks and progressively merge these chunks until only one remains (as detailed in Section \ref{chunks}).
+  This approach not only resolves the issue of defining neighbors but also mitigates the problem of premature convergence, which is common in swarm optimization algorithms.
+
+  \item For the algorithm to be robust, it must ensure that no parameter values lead to divergence.
+  As an initial solution, we introduced a \texttt{max\_speed} parameter that changes over time to ensure stability.
+  However, since this parameter was not described in the original paper, we omitted it in later implementations.
+
+  This omission led to instabilities, particularly with the enemy coefficient (if it exceeded 0.5) and the separation coefficient.
+  For example, the separation coefficient could cause exponential growth: if the distance between a dragonfly and the average position of the population is very small (e.g., 0.001) but there are many dragonflies in the neighborhood (e.g., 1000), the coefficient would become large (e.g., 1).
+  In the next iteration, the dragonflies would move far from the center, causing the distance to grow exponentially.
+
+  To address this, we modified the separation formula as follows:
+  $$S_i = - \frac{\sum_{j=1}^M (P - P_j)}{M},$$
+  where $M$ is the number of neighbors.
+  This normalization ensures that the separation coefficient remains stable regardless of the number of neighbors.
+
+  \item The enemy coefficient, as described in the original paper, does not behave as intended.
+  Specifically, the sum of vectors does not necessarily point outward from the enemy (e.g., when position vectors are aligned).
+  We assumed this was a typo, but even using the difference of vectors would not suffice.
+  If the enemy is too close to the dragonfly, the difference vector would become smaller instead of larger, leading to incorrect behavior.
+
+  A potential solution would be to normalize the difference vector by its length (i.e., use the inverse of the vector length).
+  However, this formulation would deviate significantly from the original paper.
+  As a result, we decided to retain the original formulation and evaluate its behavior empirically.
+
+\end{itemize}
+
+\subsection*{Chunk-based Optimization}
+\label{chunks}
+As mentioned earlier, the dragonfly algorithm, like many other swarm optimization solutions, has a high chance of getting stuck in local minima, limiting its ability to find better candidates.
+One common approach to mitigate this issue in swarm algorithms is to avoid relying solely on Euclidean distance for neighborhood calculations.
+Instead, the population can be divided into random groups, which are progressively merged over time until only one group remains.
+This strategy allows for greater exploration in the early stages and focuses on exploitation as the algorithm converges.
+
+In order to do that, and efficently, we divided the search space into chunks, with each thread processing an equal number of chunks.
+Every $x$ iterations, these chunks are merged into larger logical chunks, each containing a power of 2 smaller chunks.
+These logical chunks act as neighborhoods for the dragonflies.
+
+To share the contributions of each chunk, we used a modified butterfly algorithm where the contributions from all elements in a chunk is shared across all the relevant threads.
+As an optimization, if a chunk is entirely contained within a single thread, the contributions are shared locally without inter-thread communication.
+
+Figure \ref{fig:chunks} illustrates the communication process.
+The circles represent different chunks, and the colors represent logical chunks.
+The blue arrows indicate communication within a process, while the red arrows represent communication across processes.
 
 \begin{figure}[H]
-  \includegraphics[scale=0.3]{img/chunks.png}
+  \includegraphics[scale=0.21]{img/chunks.png}
   \centering
-  \caption{illustration of communication between chunks}
+  \caption{Illustration of communication between chunks.}
+  \label{fig:chunks}
 \end{figure}
 
-\noindent The implementation of this process is quite simple: each chunk will contain a subset
-of components of the global population that will specialize to compute the point with the best local
-fitness for both enemies and food of that specific region, then the local status will be communicated
-to all other chunks, updating in this way the global status of the program thanks to the update of positions,
-and the local configuration of each region; the algorithm ends displaying the best fitness found
-during the execution.
+\subsection*{Chunk-based Optimization Version 2}
 
-\section{Critical analysis of the coefficients}
-From this implementation we discovered some critical points that should be improved:
-\begin{itemize}
-  \item In the original paper\cite{Original} and in the algorithm often uses the therm neighbours without giving a precise definition.
-  Assuming a general distance metric (eg Euclidean Distance), could be an option, but later in the paper it describes how the alghorithm runs in linear time with respect to the number of dragonflies and the iteration count.
-  This would be false in case of euclidean distance, because in order to define for each butterfly the neighbours, we should calculate the distance between each pair of dragonflies, which would lead to a time complexity of $O(N^2)$.
-  Maybe using some complex geospatial algorithms it could be possible to archive that, but we doubt this was the original intention of the authors.
-  Later in the paper also another crithical problem in respect to other algorithm was the premature convergence.
-So we decided to divide the population in chunks, and progressively merge chunks until we will remain with only one.
-  So in this way we addressed both problems.
-  \item In order to be a solid algorithm it must be garanteed that does not exist some values that would make the solution diverge.
-  As a first approach we introduced a max\_speed\ parameter that would change over time and would assure stability. 
-  But given that in the original paper it was not described we omitted from later implementations.
-  This lead to instabilities from the enemy coefficient (if bigger than 0.5), but mainly from the separation coefficient.
-  This is a problem, because we could be see the coefficient as n times the distance between the dragonfly and the average position of the population. For example if the distance is 0.001, 
-  but there are 1000 dragonflies in the neighbourhood, then the coefficient would be 1, at the very next iteration the dragonflies would have a distance of 1 from the center and so on, causing an exponential growth.
-  In order to fix this we modify the separation formula as: $S_i = - \frac{\sum^M_{j=1}{P - P_j}}{M}$.
+To improve the previous algorithm, we simplified the chunk structure by removing the first chunk block abstraction.
+This modification ensures that there is only one type of chunk, eliminating an unnecessary layer of abstraction, and decreasing the risk of off by one error.
+In Figure~\ref{fig:chunks}, the circles now represent butterflies, and the colors correspond to logical chunks.
 
-  \item  The enemy coefficient does not represent what the author describes. The sum of vectors does not necessarily point outwards in respect to the enemy (eg with aligned pos vectors). 
-  We assumed a typo, but even the difference of the vectors would not be enough, because if the enemy is too close to the dragonfly, the difference would get smaller and not bigger.
-  A possible solution would be to normalize the difference vector with the inverse of the vector length, but the formulation would have been to different from the original one.
-So we decided to keep the original formulation and see if it actualy would do something.
-\end{itemize}
+Additionally, we modified the communication algorithm to address a limitation of the butterfly algorithm, which requires the number of participating threads to be a power of two.
+To overcome this constraint, we implemented a two-step approach: first, contributions are accumulated into a single element, and then the result is shared with all participating processes.
+This approach ensures compatibility with any number of threads while maintaining efficient communication.
 
@@
 
-OpenMPI is a library that enables code execution on multiple cores, wheter they belong
-to the same CPU or not; it can be useful in case of performace improvement of those algorihms
-that require to solve complex tasks, just like the dragonfly algorithm: in fact this library can be easily 
-adapted to the illustrated implementation in order to boost performance.
+The Message Passing Interface (MPI) is a portable message-passing standard designed to function on parallel computing architectures.
+For our implementation it is usefull because it let us to express multiprocess comunication across multiple machines.
 
 \subsection*{MPI on chunks}
-As we said earlier chunks are independent from each otherm but usually they need to comunicate
-with their neighbours in order to keep updated the global status of the program execution. 
-It is possible to run chunks groups simoultaneously, using MPI builtin function to connect 
-the processes.  The instructions computed by each process are available in a custom function
-called \textit{dragonfly-compute}; after calculating the ID of the initial and final chunks
-and initializing message structures using the \textbf{offsetof} macro
-the function will execute the update loop that calculates the new velocities and positions 
-of the dragonfly for each chunk associated. To do this we first need to update the current
-location of enemies and food present in the chunk using the function \textbf{computation\_status\_merge},
-using the position of food and enemies of other chunks retrieved by the function \textbf{message\_broadcast};
-finally the process will proceed to update current position and velocities of dragonfiles using 
-the function \textbf{dragonfly\_compute\_step} illustrated earlier
+By far the biggest challenge in the parallelization with MPI was the comunication of the chunks partial computations (given that a chunk may or may not span multiple process).
+Mpi has an atomic function called $sendrecv$, that permits to send a message from a thread while receving from (possibly) another thread.
+This is super useful, but it could have caused deadlocks if every process has to comunicate with multiple chunks.
+So in order to use that we needed to ensure that each process is responsible of the accumulation of one chunk.
+We did this by ensuring that if a chunk starts in another process, and ends in this one, the process comunicate its contribute to the left process.
+Then the accumulation is performed and the result is shared back to the current block.
 
-\subsection*{MPI and parameters}
-Since MPI has been useful in the parallelization of the chunks then it can also become useful in
-the training of parameters.
+\begin{figure}[H]
+  \includegraphics[scale=0.25]{img/MPI.png}
+  \centering
+  \caption{Illustration of communication between processes}
+  \label{fig:MPI}
+\end{figure}
+
+The algorithm is divided in steps (isolated by an MPI barrier):
+\begin{itemize}
+  \item \textbf{Send left}: first, the process that have the end of a block, share theyr contribute to the left
+  \item \textbf{Accumulate and share} Then it is accumulated in the first process and broadcasted back to all partecipating processes
+  \item \textbf{Send right} Then the rightmost process in the block receives the accumulated value from the left.
+\end{itemize}
+
+In this way the comunication is efficient, and looking at the results it scales very wheel. More on that in section \ref{results}.
+
+\subsection*{MPI and training}
+Since the parallelization with MPI is highly efficient, it becomes feasible to refine the hyperparameters of small problem instances by scaling the computation across hundreds of cores.
+In the results section, we demonstrate how the hyperparameters for smaller instances of the dragonfly algorithm were trained on a shifted and rotated Rastrigin function, which is commonly used as a benchmark for optimization algorithms.
+
+The learned hyperparameters were then fed back into the algorithm to enhance its performance.
+This iterative process was repeated twice, allowing the algorithm to progressively improve its ability to navigate complex search spaces.
+By leveraging the scalability of MPI, we were able to significantly reduce the time required for hyperparameter tuning while maintaining the accuracy of the results.
 
@@
 \subsection*{results}
+\label{results}
 Now it is possible to show results obtained from the cluster,

### 2025-08-27 text-paper/paper.tex
small changes

@@
 using the configuration with 4 cores and 1 thread will be considered
-equal as the serial time
-\newline
+equal as the serial time.
+The reason for this approach is that a single core computation would require 1024 times the workload of the 1024 process version, but we calibrated the 1024 core version to take 3 minutes.
 

### 2025-08-27 code/dragonfly-common.c
better approach

@@ -12,6 +12,10 @@
 // #ifdef USE_MPI
 #include <mpi.h>
 #include <unistd.h>
+
+// TODO remove signature
+void new_computation_accumulate_2(Dragonfly *d, LogicalChunk *current_chunk,
+                                  unsigned int *seed, unsigned int nr_threads);
 // #endif
 
 // n, chunks, iterations, dim
@@ -198,8 +202,6 @@ void Build_mpi_type_computation_status(MPI_Datatype *data) {
 
 // #endif
 
-
-
 unsigned int assing_logical_chunks(unsigned int phisical_chunk_start,
                                    unsigned int phisical_chunk_end,
                                    unsigned int phisical_chunk_size,
@@ -286,8 +288,6 @@ void comunicate(LogicalChunk *cur, MPI_Datatype type,
     }
   }
 }
-
-
 
 void inner_dragonfly_step(Dragonfly *d, LogicalChunk chunk,
                           unsigned int random_seed) {
@@ -392,7 +392,6 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
     Build_mpi_type_computation_status(computation_status_type);
   }
 
-
   LogicalChunk *local_chunks =
       malloc(sizeof(LogicalChunk) * maximum_logical_chunks);
 
@@ -419,7 +418,8 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
     //  1) accumulate
     for (unsigned int j = 0; j < n_local_chunks; j++) {
       // TODO paralelization opportunity if handling correctly random seed
-      new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,p.threads_per_process);
+      new_computation_accumulate_2(&cur, &local_chunks[j], &cur.seed,
+                                   p.threads_per_process);
     }
     if (threads > 1) {
 
@@ -559,35 +559,36 @@ void update_chunk_size(ChunkSize *c) {
   c->count = (unsigned int)(round(temp));
 }
 
-void inner_computation_accumulate(Dragonfly* d,unsigned int thread_start,unsigned int thread_end,
-																	float* local_cumulated_pos,float* local_cumulated_speed,
-																	float* local_food_fitness,float* local_enemy_fitness,
-																	unsigned int* local_indexes,unsigned int dim,unsigned int *local_seed){
[…]

### 2025-08-28 code/dragonfly-common.c
implemented openmp in the refractored computation_accumulate

@@ -13,11 +13,6 @@
 #include <mpi.h>
 #include <unistd.h>
 
-// TODO remove signature
-void new_computation_accumulate_2(Dragonfly *d, LogicalChunk *current_chunk,
-                                  unsigned int *seed, unsigned int nr_threads);
-// #endif
-
 // n, chunks, iterations, dim
 Parameters parameter_parse(int argc, char *argv[]) {
   Parameters p;
@@ -418,7 +413,7 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
     //  1) accumulate
     for (unsigned int j = 0; j < n_local_chunks; j++) {
       // TODO paralelization opportunity if handling correctly random seed
-      new_computation_accumulate_2(&cur, &local_chunks[j], &cur.seed,
+      computation_accumulate(&cur, &local_chunks[j], &cur.seed,
                                    p.threads_per_process);
     }
     if (threads > 1) {
@@ -559,45 +554,7 @@ void update_chunk_size(ChunkSize *c) {
   c->count = (unsigned int)(round(temp));
 }
 
-void inner_computation_accumulate(
-    Dragonfly *d, unsigned int thread_start, unsigned int thread_end,
-    float *local_cumulated_pos, float *local_cumulated_speed,
-    float *local_food_fitness, float *local_enemy_fitness,
-    unsigned int *local_indexes, unsigned int dim, unsigned int *local_seed) {
-  for (unsigned int k = thread_start; k < thread_end; k++) {
-    float *iter_pos = d->positions + dim * k;
-    float *iter_speed = d->speeds + dim * k;
-    sum_assign(local_cumulated_pos, iter_pos, dim);
-    sum_assign(local_cumulated_speed, iter_speed, dim);
-
-    float fitness = d->fitness(iter_pos, local_seed, dim);
-
-    if (fitness > *local_food_fitness) {
-      local_indexes[0] = k;
-      *local_food_fitness = fitness;
-    }
-    if (fitness < *local_enemy_fitness) {
-      local_indexes[1] = k;
-      *local_enemy_fitness = fitness;
-    }
-  }
-}
-
-void update_status(Dragonfly *d, LogicalChunk *current_chunk,
-                   float next_food_fitness, float next_enemy_fitness,
-                   unsigned int *indexes, unsigned int start, unsigned int end,
-                   unsigned int dim) {
-
[…]

### 2025-08-28 code/dragonfly-common.h

@@ -106,4 +106,10 @@ void computation_status_merge(ComputationStatus *out, ComputationStatus *in,
 float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
                          unsigned int threads, unsigned int rank_id,
                          float space_size, unsigned int srand);
+
+void inner_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
+                                      unsigned int *seed, unsigned start,
+                                      unsigned end);
+void computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
+                                  unsigned int *seed, unsigned int nr_threads);
 #endif

### 2025-08-28 code/bin_mpich.c
better rotated fitness handling

@@ -3,13 +3,12 @@
 #include <stddef.h>
 #include <stdio.h>
 #include <stdlib.h>
-#include <string.h>
 #include <time.h>
 #include <unistd.h>
 
 #include "dragonfly-common.h"
 #include "utils.h"
-
+unsigned counter=0;
 int main(int argc, char *argv[]) {
   MPI_Init(NULL, NULL);
   // wait for all the process to start
@@ -56,11 +55,11 @@ float wi[14] = {
   MPI_Comm_rank(MPI_COMM_WORLD, &rank);
   Fitness fitness = rastrigin_fitness;
 
-  float *res = dragonfly_compute(p, w, c, fitness, comm_size, rank, 100.0, start_time);
+  float *res = dragonfly_compute(p, w, c, fitness, NULL, 0, comm_size, rank, 100.0, start_time);
 
   MPI_Barrier(MPI_COMM_WORLD);
   unsigned int s = 0;
-  float fit = fitness(res, &s, p.problem_dimensions);
+  float fit = fitness(res, &s, p.problem_dimensions, NULL);
 
   if (rank == 0) {
     printf("found fitness=%f\n", fit);
@@ -71,5 +70,7 @@ float wi[14] = {
     printf("Execution time = %f\n", duration);
   }
   free(res);
+  printf("%d\n", counter);
   MPI_Finalize();
+  
 }

### 2025-08-28 code/bin_training_mpich.c

@@ -8,7 +8,14 @@
 #include <time.h>
 #include <unistd.h>
 
-float eval(float *wi, unsigned int *seedi, unsigned int d) {
+float eval(float *wi, unsigned int *seedi, unsigned int d, void *data) {
+  (void)data;
+  if (d < 14) {
+    fprintf(stderr,
+            "impossible to compute with %d dimensions it must be at least 14\n",
+            d);
+    exit(-1);
+  }
   unsigned int seed = *seedi;
   for (int i = 0; i < 14; i++) {
     if (wi[i] < 0.0) {
@@ -31,7 +38,7 @@ float eval(float *wi, unsigned int *seedi, unsigned int d) {
     wi[10] /= 1.2 * m;
     wi[11] /= 1.2 * m;
   }
-  (void)d;
+
   Weights w = {
       // exploring
       .al = {wi[0], wi[1]},
@@ -49,31 +56,25 @@ float eval(float *wi, unsigned int *seedi, unsigned int d) {
                   .iterations = 80,
                   .threads_per_process = 1};
   ChunkSize c = new_chunk_size(p.starting_chunk_count, 1, p.iterations);
-  Fitness fitness = shifted_fitness;
+  Fitness fitness = rastrigin_fitness;
   float avg = 0.0;
   int N = 30;
   for (int i = 0; i < N; i++) {
 
-    float *shifted_tmp = malloc(sizeof(float) * p.problem_dimensions);
-    float *shifted_rotation =
-        malloc(sizeof(float) * p.problem_dimensions * p.problem_dimensions);
-    float *shifted_shift = init_array(p.problem_dimensions, 80.0, &seed);
-    init_matrix(shifted_rotation, 90.0, p.problem_dimensions, &seed);
+    void *shifted_fitness_data = malloc_shifted_fitness(fitness, 80.0, 90.0, &seed,
+                                                   p.problem_dimensions);
 
-    init_shifted_fitness(shifted_tmp, shifted_rotation, shifted_shift,
-                         rastrigin_fitness);
+    float *res =
+        dragonfly_compute(p, w, c, shifted_fitness, shifted_fitness_data,
+                          sizeof(ShiftedFitnessParams), 1, 0, 100.0, seed);
 
-    float *res = dragonfly_compute(p, w, c, fitness, 1, 0, 100.0, seed);
-
-    avg += fitness(res, &seed, p.problem_dimensions);
[…]

### 2025-08-28 code/dragonfly-common.c

@@ -1,3 +1,4 @@
+#define USE_OPENMP
 #include "dragonfly-common.h"
 #include "stdlib.h"
 #include "string.h"
@@ -14,8 +15,8 @@
 #include <unistd.h>
 
 // TODO remove signature
-void new_computation_accumulate_2(Dragonfly *d, LogicalChunk *current_chunk,
-                                  unsigned int *seed, unsigned int nr_threads);
+void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
+                                unsigned int *seed, unsigned int nr_threads);
 // #endif
 
 // n, chunks, iterations, dim
@@ -88,6 +89,7 @@ void weights_step(Weights *w) {
 Dragonfly dragonfly_new(unsigned int dimensions, unsigned int start,
                         unsigned int end, unsigned int iterations,
                         float space_size, Weights weights, Fitness fitness,
+                        void *fitness_data, unsigned fitness_data_size,
                         unsigned int random_seed) {
   // computes weigths progression
 
@@ -101,6 +103,8 @@ Dragonfly dragonfly_new(unsigned int dimensions, unsigned int start,
       .space_size = space_size,
       .w = weights,
       .fitness = fitness,
+      .fitness_data = fitness_data,
+      .fitness_data_size = fitness_data_size,
       .seed = random_seed,
   };
   return d;
@@ -153,35 +157,6 @@ void computation_status_merge(ComputationStatus *out, ComputationStatus *in,
   }
   out->n += in->n;
 }
-
-// #ifdef USE_MPI
-/*
-void raw_send_recv(Message *send, unsigned int destination, Message
-*recv_buffer, unsigned int source, MPI_Datatype *data_type) { MPI_Sendrecv(send,
-1, *data_type, destination, 0, recv_buffer, 1, *data_type, source, 0,
-MPI_COMM_WORLD, MPI_STATUS_IGNORE);
-}
-
-
-void message_broadcast(Message *message, unsigned int index, int n, MPI_Datatype
-*data_type){
-
-  Message recv_buffer;
-  for(unsigned int steps=0; (1<<steps)<n; steps++){
-    unsigned int index_other;
-    if ((index>>steps) % 2 == 0) {
-      index_other = index+(1<<steps);
-    } else {
-      index_other = index-(1<<steps);
-    }
-    raw_send_recv(message, index_other, &recv_buffer, index_other, data_type);
-    memcpy(&message->status[recv_buffer.start_chunk],
[…]

### 2025-08-28 code/dragonfly-common.h

@@ -31,7 +31,7 @@ void update_chunk_size(ChunkSize *c);
 
 // Newtype for Fitness functions, they should take the input vector, one int as
 // random number generator seed, and the dimension count
-typedef float (*Fitness)(float *, unsigned int *, unsigned int);
+typedef float (*Fitness)(float *, unsigned int *, unsigned int, void *data);
 
 // function used to compute the step for the linear progression of the weights.
 void weights_compute_steps(Weights *w, unsigned int steps);
@@ -46,6 +46,10 @@ typedef struct {
   float space_size;
   // fitness function
   Fitness fitness;
+  // fitness additional data
+  void *fitness_data;
+  unsigned fitness_data_size;
+
   // adaptive weights
   Weights w;
 
@@ -64,7 +68,7 @@ typedef struct {
 
 Dragonfly dragonfly_new(unsigned int dimensions, unsigned int start,
                         unsigned int end, unsigned int iterations,
-                        float space_size, Weights weights, Fitness fitness,
+                        float space_size, Weights weights, Fitness fitness, void *fitness_data, unsigned fitness_data_size,
                         unsigned int random_seed);
 
 void dragonfly_alloc(Dragonfly *d);
@@ -104,6 +108,7 @@ Parameters parameter_parse(int argc, char *argv[]);
 void computation_status_merge(ComputationStatus *out, ComputationStatus *in,
                               unsigned int dim);
 float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
-                         unsigned int threads, unsigned int rank_id,
+                         void *fitness_data, unsigned fitness_data_size,
+                         unsigned int process_count, unsigned int rank_id,
                          float space_size, unsigned int srand);
 #endif

### 2025-08-28 code/utils.c

@@ -22,8 +22,6 @@ void* my_malloc(size_t size, const char *file, int line, const char *func)
 }*/
 #include "utils.h"
 
-
-
 float *init_array(unsigned int dimensions, float range_max,
                   unsigned int *seed) {
   float *arr = malloc(dimensions * sizeof(float));
@@ -120,7 +118,8 @@ void brownian_motion(float *inp, unsigned int dim, unsigned int *seed) {
   }
 }
 
-float rastrigin_fitness(float *inp, unsigned int *seed, unsigned int dim) {
+float rastrigin_fitness(float *inp, unsigned int *seed, unsigned int dim, void *data) {
+  (void)(data);
   (void)(seed);
   float ret = 10.0 * dim;
   for (unsigned int i = 0; i < dim; i++) {
@@ -129,7 +128,8 @@ float rastrigin_fitness(float *inp, unsigned int *seed, unsigned int dim) {
   return -ret;
 }
 
-float sphere_fitness(float *inp, unsigned int *seed, unsigned int dim) {
+float sphere_fitness(float *inp, unsigned int *seed, unsigned int dim, void *data) {
+  (void)(data);
   (void)(seed);
   float res = 0.0;
   for (unsigned int i = 0; i < dim; i++) {
@@ -137,7 +137,8 @@ float sphere_fitness(float *inp, unsigned int *seed, unsigned int dim) {
   }
   return -res;
 }
-float rosenblock_fitness(float *inp, unsigned int *seed, unsigned int dim) {
+float rosenblock_fitness(float *inp, unsigned int *seed, unsigned int dim, void *data) {
+  (void)(data);
   (void)(seed);
   float res = 0.0;
   for (unsigned int i = 0; i < dim - 1; i++) {
@@ -147,23 +148,25 @@ float rosenblock_fitness(float *inp, unsigned int *seed, unsigned int dim) {
   return -res;
 }
 
-float *ROTATION;
-float *SHIFT;
-Fitness FITNESS;
-float *TMP;
-void init_shifted_fitness(float *tmp, float *rotation, float *shift,
-                          Fitness fitness) {
-  ROTATION = rotation;
-  SHIFT = shift;
-  FITNESS = fitness;
-  TMP = tmp;
+ShiftedFitnessParams *malloc_shifted_fitness(Fitness fitness, float shift_range,
+                                             float rotation_range,
[…]

### 2025-08-28 code/utils.h

@@ -21,21 +21,27 @@ void zeroed(float *dest, unsigned int size);
 
 void brownian_motion(float* inp, unsigned int dim, unsigned int * seed);
 // fitness functions
-float rastrigin_fitness(float *inp, unsigned int *, unsigned int dim);
-float sphere_fitness(float *inp, unsigned int *, unsigned int dim);
-float rosenblock_fitness(float *inp, unsigned int*, unsigned int dim);
+float rastrigin_fitness(float *inp, unsigned int *, unsigned int dim, void *data);
+float sphere_fitness(float *inp, unsigned int *, unsigned int dim, void *data);
+float rosenblock_fitness(float *inp, unsigned int*, unsigned int dim, void *data);
 
-float shifted_fitness(float *inp, unsigned int* seed, unsigned int dim);
-void init_shifted_fitness(float *tmp, float * rotation, float * shift, Fitness fitness);
 void init_matrix(float *inp, float range_max, unsigned int dim, unsigned int *seed);
 
+#define MAX_PROBLEM_DIMENSIONS 16
+typedef struct {
+  Fitness fitness;
+  float tmp[MAX_PROBLEM_DIMENSIONS];
+  float rotation[MAX_PROBLEM_DIMENSIONS*MAX_PROBLEM_DIMENSIONS];
+  float shift[MAX_PROBLEM_DIMENSIONS];
 
-void dragonfly_compute_step(Dragonfly *d, float *average_speed,
-    float *cumulated_pos, float *food, float *enemy,
-    unsigned int N, unsigned int NUM_THREADS);
-void new_computation_accumulate(Dragonfly *d, 
-																LogicalChunk *current_chunk,
-                                unsigned int *seed,
-																unsigned int nr_threads);
+} ShiftedFitnessParams;
+
+ShiftedFitnessParams *malloc_shifted_fitness(Fitness fitness, float shift_range,
+                                             float rotation_range,
+                                             unsigned *seed, unsigned dim);
+void free_shifted_fitness(ShiftedFitnessParams *s);
+
+float shifted_fitness(float *inp, unsigned int *seed, unsigned int dim,
+                      void *data);
 
 #endif

### 2025-08-28 code/dragonfly-common.c
fixed openmp error

@@ -1,4 +1,4 @@
-#define USE_OPENMP
+#include <float.h>
 #include "dragonfly-common.h"
 #include "stdlib.h"
 #include "string.h"
@@ -9,8 +9,6 @@
 #include <stddef.h>
 #include <stdio.h>
 
-#include <time.h>
-// #ifdef USE_MPI
 #include <mpi.h>
 #include <unistd.h>
 
@@ -393,10 +391,23 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
         cur.start, cur.end, phisical_chunk_size, p.population_size,
         process_count - rank_id - 1, logical_chunk_size, local_chunks);
     //  1) accumulate
-    for (unsigned int j = 0; j < n_local_chunks; j++) {
-      // TODO paralelization opportunity if handling correctly random seed
-      new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
-                                 p.threads_per_process);
+    if (p.threads_per_process > 1) {
+      #ifdef USE_OPENMP
+#pragma omp parallel for num_threads(p.threads_per_process)
+      for (unsigned int j = 0; j < n_local_chunks; j++) {
+        new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
+                                   p.threads_per_process);
+      }
+      #else
+      fprintf(stderr, "compiled without openmp, code unavailable\n");
+      exit(-1);
+      #endif
+
+    } else {
+      for (unsigned int j = 0; j < n_local_chunks; j++) {
+        new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
+                                   p.threads_per_process);
+      }
     }
     if (process_count > 1) {
 
@@ -560,9 +571,8 @@ void inner_new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
   zeroed(cumulated_pos, dim);
   zeroed(cumulated_speed, dim);
 
-  float next_enemy_fitness = d->fitness(d->positions+(start-d->start)*dim, seed, dim, fitness_data);
-  float next_food_fitness = next_enemy_fitness;
-  // status->n = d->local_n;
+  float next_enemy_fitness = FLT_MAX;
+  float next_food_fitness = FLT_MIN;
 
   unsigned int indexes[2] = {0, 0};
   end = end - d->start;
@@ -608,9 +618,9 @@ void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
   LogicalChunk *temp_chunks =
[…]

### 2025-08-28 code/dragonfly-common.c
fix bug

@@ -1,8 +1,8 @@
-#include <float.h>
 #include "dragonfly-common.h"
 #include "stdlib.h"
 #include "string.h"
 #include "utils.h"
+#include <float.h>
 
 #include <assert.h>
 #include <stdbool.h>
@@ -304,7 +304,14 @@ void inner_dragonfly_step(Dragonfly *d, LogicalChunk chunk,
       cur_speed[i] += d->w.e * E;
       cur_speed[i] += levy;
 
+      cur_speed[i] = min(cur_speed[i], d->space_size);
+      cur_speed[i] = max(cur_speed[i], -d->space_size);
+
       cur_pos[i] += cur_speed[i];
+
+      cur_pos[i] = min(cur_pos[i], d->space_size);
+      cur_pos[i] = max(cur_pos[i], -d->space_size);
+      
     }
   }
 }
@@ -391,23 +398,12 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
         cur.start, cur.end, phisical_chunk_size, p.population_size,
         process_count - rank_id - 1, logical_chunk_size, local_chunks);
     //  1) accumulate
-    if (p.threads_per_process > 1) {
-      #ifdef USE_OPENMP
-#pragma omp parallel for num_threads(p.threads_per_process)
-      for (unsigned int j = 0; j < n_local_chunks; j++) {
-        new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
-                                   p.threads_per_process);
-      }
-      #else
-      fprintf(stderr, "compiled without openmp, code unavailable\n");
-      exit(-1);
-      #endif
-
-    } else {
-      for (unsigned int j = 0; j < n_local_chunks; j++) {
-        new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
-                                   p.threads_per_process);
-      }
+    #ifdef USE_OPENMP
+    #pragma omp parallel for num_threads(p.threads_per_process)
+    #endif
+    for (unsigned int j = 0; j < n_local_chunks; j++) {
+      new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
+                                 p.threads_per_process);
     }
     if (process_count > 1) {
 
@@ -572,8 +568,7 @@ void inner_new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
   zeroed(cumulated_speed, dim);
 
[…]

### 2025-08-28 code/dragonfly-common.c
fix bug

@@ -614,9 +614,10 @@ void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
       (LogicalChunk *)malloc(sizeof(LogicalChunk) * nr_threads);
   unsigned int elems_per_thread = (end - start) / nr_threads;
 
-#pragma omp parallel num_threads(nr_threads)
+#pragma omp parallel for num_threads(nr_threads)
+for(unsigned int thread_id=0; thread_id<nr_threads; thread_id++)
   {
-    unsigned int thread_id = omp_get_thread_num();
+    //unsigned int thread_id = omp_get_thread_num();
     void *tmp_fitness_data = NULL;
     if (d->fitness_data_size != 0) {
       tmp_fitness_data = malloc(d->fitness_data_size);
@@ -638,7 +639,7 @@ void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
     }
   }
   memcpy(current_chunk, &temp_chunks[0], sizeof(LogicalChunk));
-  for (unsigned int i = 0; i < nr_threads; i++) {
+  for (unsigned int i = 1; i < nr_threads; i++) {
     computation_status_merge(&current_chunk->comp,
                              &temp_chunks[i].comp, dim);
   }

### 2025-08-28 code/bin_training_mpich.c
fix bug

@@ -84,8 +84,8 @@ int main(int argc, char **argv) {
   MPI_Init(NULL, NULL);
   // wait for all the process to start
   MPI_Barrier(MPI_COMM_WORLD);
-  clock_t start_time;
-  start_time = clock();
+  struct timespec start_time, end_time;
+    clock_gettime(CLOCK_MONOTONIC, &start_time);
   int comm_size, rank;
   MPI_Comm_size(MPI_COMM_WORLD, &comm_size);
   MPI_Comm_rank(MPI_COMM_WORLD, &rank);
@@ -109,7 +109,7 @@ int main(int argc, char **argv) {
       .ll = {wi[12], wi[13]},
   };
   float *res = dragonfly_compute(p, w, c, fitness, NULL, 0, comm_size, rank,
-                                 2.0, start_time + rank);
+                                 2.0, start_time.tv_nsec + rank);
   // float *res = malloc(sizeof(float)*p.population_size);
   MPI_Barrier(MPI_COMM_WORLD);
 
@@ -120,7 +120,8 @@ int main(int argc, char **argv) {
     for (unsigned int i = 0; i < p.problem_dimensions; i++) {
       printf("%f\n", res[i]);
     }
-    double duration = (double)(clock() - start_time) / CLOCKS_PER_SEC;
+    clock_gettime(CLOCK_MONOTONIC, &end_time);
+    double duration = (double)(end_time.tv_nsec - start_time.tv_nsec) / 1e9 + (end_time.tv_sec - start_time.tv_sec);
     printf("Execution time = %f\n", duration);
   }
   free(res);

### 2025-08-28 code/dragonfly-common.c

@@ -2,6 +2,7 @@
 #include "stdlib.h"
 #include "string.h"
 #include "utils.h"
+#include <time.h>
 #include <float.h>
 
 #include <assert.h>
@@ -311,7 +312,6 @@ void inner_dragonfly_step(Dragonfly *d, LogicalChunk chunk,
 
       cur_pos[i] = min(cur_pos[i], d->space_size);
       cur_pos[i] = max(cur_pos[i], -d->space_size);
-      
     }
   }
 }
@@ -398,13 +398,24 @@ float *dragonfly_compute(Parameters p, Weights w, ChunkSize c, Fitness fitness,
         cur.start, cur.end, phisical_chunk_size, p.population_size,
         process_count - rank_id - 1, logical_chunk_size, local_chunks);
     //  1) accumulate
-    #ifdef USE_OPENMP
-    #pragma omp parallel for num_threads(p.threads_per_process)
-    #endif
-    for (unsigned int j = 0; j < n_local_chunks; j++) {
-      new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
-                                 p.threads_per_process);
+
+    //accumulate time with wall-clock time (not CPU time)
+
+    if (p.threads_per_process > 1) {
+#pragma omp parallel for num_threads(p.threads_per_process)
+      for (unsigned int j = 0; j < n_local_chunks; j++) {
+
+        new_computation_accumulate(&cur, &local_chunks[j], &cur.seed,
+                                   p.threads_per_process);
+      }
+    } else
+
+    {
+      for (unsigned int j = 0; j < n_local_chunks; j++) {
+        new_computation_accumulate(&cur, &local_chunks[j], &cur.seed, 1);
+      }
     }
+
     if (process_count > 1) {
 
       // 2) send rightmost part to 1 left

### 2025-08-29 text-paper/paper.tex
openmp section

@@
 \begin{itemize}
-  \item \textbf{Send left}: first, the process that have the end of a block, share theyr contribute to the left
+  \item \textbf{Send left}: first, the process that have the end of a block, share their contribute to the left
   \item \textbf{Accumulate and share} Then it is accumulated in the first process and broadcasted back to all partecipating processes
@@
 
-\section{OpenMP implementation}
+\section{OpenMP optimization}
 
 OpenMP is a low level API that allows us to execute a sequence of instructions on different threads belonging
-to the same CPU core, improving in this way the performances of our implementation. This feature can be used to 
-speed up the execution of some slow functions, in our case these will be dragonfly-compute-step and message-accumulate
+to the same CPU core, improving in this way the performances of our implementation.
+Thanks to a compiler that supports this technology, in our case it is \textit{gcc},
+we can include  pragma directives in our code,
+which will wrap the set of instructions that we are interested to execute
+in parallel. This feature can be used to 
+speed up the execution of some slow functions like the \textit{computation-accumulate}
 
-\subsection*{General use}
+\subsection*{computation\_accumulate}
 
-First of all it is necessary to understand how OpenMP should be inserted in the code: first of all we have to use 
-a compiler that supports this technology, in our case we have gcc, which allows OpenMP execution thanks to the flag -fopenmp.
-Second we need to include a pragma directive in our code, which will wrap the set of instructions that we are interested to execute
-in parallel. Since the implemented results will perform loops mainly then it is fundamental to calculate the work balance between the number of iterations, if the main
-body is a loop,
-and the number of threads available. Each thread has to define the starting and the end point of the loop that has to execute,
-calculating the starting index as $rank*ratio$ and the end one as $ratio*(rank+1)$, where $rank$ is the ID of the thread
+This function, after defining the starting and the ending points, compute
+the cumulated sums of dragonfiles positions and speeds 
+and saves the indexes of dragonfiles positions with the best fitness, using them
+to set the next locations of food and enemies.
 
-\subsection*{dragonfly-compute-step}
+\begin{lstlisting}[language=C,caption={the serial version of \textit{computation\_accumulate}}]
+  unsigned dim = d->dim;
+  float *cumulated_pos = current_chunk->comp.cumulated_pos;
+  float *cumulated_speed = current_chunk->comp.cumulated_speeds;
+  zeroed(cumulated_pos, dim);
+  zeroed(cumulated_speed, dim);
 
-One of the two functions that we are interested to execute using OpenMP is \textit{dragonfly-compute-step}:
-if we take a look on the serial version we notice that each computation step is independent
-from other iterations. This means that the execution flow can be split in many subtasks assigned
-to threads. The following code snippet illustrate how all the procedure has been implemented
+  float next_enemy_fitness = FLT_MAX;
+  float next_food_fitness = FLT_MIN;
+  unsigned int indexes[2] = {0, 0};
+  end = end - d->start;
 
-\begin{lstlisting}[language=C,caption={parallelized dragonfly-compute-step}]
-void dragonfly_compute_step(Dragonfly *d, 
-                            float *average_speed,
-                            float *cumulated_pos, 
-                            float *food, 
-                            float *enemy, 
-                            unsigned int N, 
-                            unsigned int NR_THREADS) {
+  for (unsigned int k = start - d->start; k < end; k++) {
+    float *iter_pos = d->positions + dim * k;
+    float *iter_speed = d->speeds + dim * k;
+    sum_assign(cumulated_pos, iter_pos, dim);
+    sum_assign(cumulated_speed, iter_speed, dim);
 
-  unsigned int base_random = rand_r(&d->seed);
-  unsigned int rest = d->N % NR_THREADS;
-  unsigned int ratio = d->N / NR_THREADS;
-
-  ...
-
-  #pragma omp parallel num_threads(NR_THREADS)
-    {
-
-      unsigned int rank = omp_get_thread_num();
-      //printf("rank=%d\n", rank);
-      unsigned int base = ratio * rank;
-      unsigned int limit = ratio * (rank + 1);
-      inner_dragonfly_step(d, average_speed, 
-                          cumulated_pos, food, 
-                          enemy, N,
-                          base, limit, base_random);
+    float fitness = d->fitness(iter_pos, 
+                               seed, dim, 
+                               fitness_data);
+    if (fitness > next_food_fitness) {
+      indexes[0] = k;
+      next_food_fitness = fitness;
     }
-
-  ...
-
-  if (rest != 0) {
-      unsigned int r_base = ratio * NR_THREADS;
-      unsigned int r_end = d->N;
-
-      inner_dragonfly_step(d, average_speed, 
-                           cumulated_pos, 
-                           food, enemy, N,
-                           r_base, r_end, base_random);
+    if (fitness < next_enemy_fitness) {
+      indexes[1] = k;
+      next_enemy_fitness = fitness;
     }
@@
 
-  weights_step(&d->w);
+  memcpy(current_chunk->comp.next_food, 
+        d->positions + indexes[0] * dim,
+        sizeof(float) * dim);
+  memcpy(current_chunk->comp.next_enemy, 
+        d->positions + indexes[1] * dim,
+        sizeof(float) * dim);
 
+  current_chunk->comp.next_enemy_fitness = next_enemy_fitness;
+  current_chunk->comp.next_food_fitness = next_food_fitness;
+  current_chunk->comp.n = end - (start - d->start);
+\end{lstlisting}
+
+The implementation of a parallelized approach in this case is obstacled by the fact that 
+threads should read and write on the same memeory areas like the \textit{cumulated\_pos},
+\textit{cumulated\_speed}, which could be 
+difficult to execute independently from each other due to race conditions. 
+In order to avoid this problem we split the current logical chunk into small ones, each
+one of them keeps track of: 
+represents:
+\begin{itemize}
+\item the local cumulated positions
+\item the local cumulated speeds
+\item the best local enemy and food fitnesses
+\item the indexes of the positions of dragonfiles with the best fitnesses
+\end{itemize}
+\noindent parallel executions takes care of a particular subchunk,
+defining also the initial and final positions by
+calculating the starting index as $thread\_ID*elems\_per\_thread$ and the end one as $start\_index+ratio$; 
+the final one however could be responsible for more dragonfiles 
+than others since the number of threads does not always
+divide the number of elements perfectly. The single threads then proceed to 
+calculate the cumulated sums and the fitnesses of the current dragonfly, 
+saving its position index if it maximize the food fitness or minimize the 
+enemy fitness. 
+\begin{lstlisting}[language=C,caption={the parallel version of \textit{computation\_accumulate}}]
+  LogicalChunk *temp_chunks =
+      (LogicalChunk *)malloc(sizeof(LogicalChunk) * nr_threads);
+  unsigned int elems_per_thread = (end - start)/nr_threads;
+
+  #pragma omp parallel num_threads(nr_threads)
+    {
+      unsigned int thread_id = omp_get_thread_num();
+      void *tmp_fitness_data = NULL;
+      if (d->fitness_data_size != 0) {
+        tmp_fitness_data = malloc(d->fitness_data_size);
+        memcpy(tmp_fitness_data, 
+               d->fitness_data, 
+               d->fitness_data_size);
+      }
+
+      memcpy(&temp_chunks[thread_id], current_chunk, sizeof(LogicalChunk));
+      unsigned int thread_start = start + (thread_id * elems_per_thread);
+      unsigned int thread_end = thread_id == (nr_threads - 1) ? end: min(thread_start + elems_per_thread,end);
+      unsigned int local_seed = 0;
+
+      inner_new_computation_accumulate(
+        d,
+        &temp_chunks[thread_id],
+        &local_seed,
+        thread_start, 
+        thread_end,tmp_fitness_data
+      );
+      if (d->fitness_data_size != 0) {
+        free(tmp_fitness_data);
+      }
     }
@@
  \end{lstlisting}
+ After the join of all threads, the program performs a reduce of subchunks, calculating
+the optimal . This methodology can be seen as a unique critical section
+executed only once after all the parallel processings have ended; then as final step the position and fitnesses
+of food and enemies is updated, without the need to define critical sections which would force the processor 
+to update the position buffers sequentially.
 
-\noindent After we got all basic informations the \textit{inner\_dragonfly\_step}
-function can be executed, wrapping it into the \textbf{pragma} directive that we have seen earlier: in this way each thread will 
-perform the necessary operations only on a specific portion of the overall positions and speeds arrays.
-If the number of threads does not divide perfectly the size of the arrays then we compute the rest
-returned from the ratio calculation and the remaining cells will be processed using the serial approach.
+\begin{lstlisting}[language=C,caption={the parallel version of \textit{computation\_accumulate}}]
 
-\subsection*{computation\_accumulate}
+  memcpy(current_chunk, 
+         &temp_chunks[0], 
+         sizeof(LogicalChunk));
 
-The \textit{computation\_accumulate}  must compute
-the cumulated positions and speeds of the chunk and find the best food and enemy positions.
-The implementation of a parallelized approach in this case  is more complicated to achieve, 
-because threads should read and write on the same memeory areas which could be 
-difficult to execute independently from each other. This time it is necessary to define
-local buffers for each thread, so they can store temporary results without the risk of using
-overlapping memeory locations. 
-\begin{figure}[H]
-  \includegraphics[scale=0.2]{img/reduce-example.png}
-  \centering
-  \caption{Illustration of the reduce process performed by the computation\_accumulate function}
-\end{figure}
-\noindent The parallel section provides that each thread must compute the local
-cumulated sums of velocities and positions, then they must select the position with the best fitness value
-in order to find the best local position of food and enemies. Once the local execution is completed
-it will update the global status of the chunk using a critical section, with one thread that can execute that
-specific portion of the program
+  for (unsigned int i = 1; i < nr_threads; i++) {
+    for (unsigned int k = 0; k < dim; k++) {
+      current_chunk->comp.cumulated_pos[k] +=
+          temp_chunks[i].comp.cumulated_pos[k];
+      current_chunk->comp.cumulated_speeds[k] +=
+          temp_chunks[i].comp.cumulated_speeds[k];
+    }
+
+    if (current_chunk->comp.next_food_fitness <
+        temp_chunks[i].comp.next_food_fitness) {
+
+          current_chunk->comp.next_food_fitness =
+          temp_chunks[i].comp.next_food_fitness;
+
+          memcpy(current_chunk->comp.next_food, 
+                  temp_chunks[i].comp.next_food,
+                  sizeof(float) * dim);
+    }
+
+    if (current_chunk->comp.next_enemy_fitness >
+        temp_chunks[i].comp.next_enemy_fitness) {
+
+        current_chunk->comp.next_enemy_fitness =
+          temp_chunks[i].comp.next_enemy_fitness;
+
+        memcpy(current_chunk->comp.next_enemy, 
+               temp_chunks[i].comp.next_enemy,
+               sizeof(float) * dim);
+    }
+  }
+  free(temp_chunks);
+  \end{lstlisting}
 

### 2025-08-29 code/dragonfly-common.c
cleanup

@@ -2,7 +2,6 @@
 #include "stdlib.h"
 #include "string.h"
 #include "utils.h"
-#include <time.h>
 #include <float.h>
 
 #include <assert.h>
@@ -12,11 +11,6 @@
 
 #include <mpi.h>
 #include <unistd.h>
-
-// TODO remove signature
-void new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
-                                unsigned int *seed, unsigned int nr_threads);
-// #endif
 
 // n, chunks, iterations, dim
 Parameters parameter_parse(int argc, char *argv[]) {
@@ -316,6 +310,100 @@ void inner_dragonfly_step(Dragonfly *d, LogicalChunk chunk,
   }
 }
 
+void inner_new_computation_accumulate(Dragonfly *d, LogicalChunk *current_chunk,
+                                      unsigned int *seed, unsigned start,
+                                      unsigned end, void *fitness_data) {
+  unsigned dim = d->dim;
+  float *cumulated_pos = current_chunk->comp.cumulated_pos;
+  float *cumulated_speed = current_chunk->comp.cumulated_speeds;
+  zeroed(cumulated_pos, dim);
+  zeroed(cumulated_speed, dim);
+
+  float next_enemy_fitness = FLT_MAX;
+  float next_food_fitness = -FLT_MAX;
+  unsigned int indexes[2] = {0, 0};
+  end = end - d->start;
+
+  for (unsigned int k = start - d->start; k < end; k++) {
+    float *iter_pos = d->positions + dim * k;
+    float *iter_speed = d->speeds + dim * k;
+    sum_assign(cumulated_pos, iter_pos, dim);
+    sum_assign(cumulated_speed, iter_speed, dim);
+
+    float fitness = d->fitness(iter_pos, seed, dim, fitness_data);
+
+    if (fitness > next_food_fitness) {
+      indexes[0] = k;
+      next_food_fitness = fitness;
+    }
+    if (fitness < next_enemy_fitness) {
+      indexes[1] = k;
+      next_enemy_fitness = fitness;
+    }
+  }
+
+  memcpy(current_chunk->comp.next_food, d->positions + indexes[0] * dim,
+         sizeof(float) * dim);
+  memcpy(current_chunk->comp.next_enemy, d->positions + indexes[1] * dim,
+         sizeof(float) * dim);
+  current_chunk->comp.next_enemy_fitness = next_enemy_fitness;
[…]

### 2025-08-29 text-paper/paper.tex
paper

@@
 \usepackage{longtable}
+\usepackage{tabularx}
 \usepackage{cite}
@@
 Swarm intelligence algorithms are inspired by the collective behavior of biological systems, where simple rules followed by individuals lead to the emergence of complex and adaptive group behavior.
-The concept was first introduced in Reynolds' Boids model (1987) \cite{TODO}, which simulated flocking behavior, and gained significant traction in Computer Science with the Ant Colony Optimization algorithm (1995) \cite{TODO}.
+The concept was first introduced in Reynolds' Boids model (1987) \cite{Boids}, which simulated flocking behavior, and gained significant traction in Computer Science with the Ant Colony Optimization algorithm (1995) \cite{Ant_Colony_optimization}.
 These algorithms are meta-heuristic in nature, leveraging decentralized and parallel exploration-exploitation strategies to solve optimization problems.
 
-The dragonfly algorithm, first proposed in 2014 and published in 2015 \cite{DAReview}, is a bio-inspired optimization algorithm that simulates the behavior of dragonfly swarms.
+The dragonfly algorithm, first proposed in 2014 and published in 2015 \cite{Original}, is a bio-inspired optimization algorithm that simulates the behavior of dragonfly swarms.
 By mimicking natural behaviors such as alignment, separation, cohesion, attraction to food sources, and distraction from enemies, the algorithm achieves a balance between exploration and exploitation.
 This makes it particularly effective at navigating complex search spaces while reducing the likelihood of being trapped in local optima.
-As of 2019, the algorithm has been cited in over 300 scientific papers, highlighting its impact and popularity in the field.
+As of 2019 \cite{Meraihi2020}, the algorithm has been cited in over 300 scientific papers, highlighting its impact and popularity in the field.
 
@@
 This optimization reduces execution time by performing all calculations locally within the loop, eliminating the need for auxiliary arrays to store intermediate coefficients.
-The rationale for dividing the separation coefficient by the number of neighbors will be explained later in this paper (TODO reference).
+The rationale for dividing the separation coefficient by the number of neighbors will be explained later in this paper in section \ref{chunks}.
 
@@
 execute multiple istances of the same problem
-\subsection*{structure of the execution}
+\subsection*{Structure of the execution}
 
@@
 resources in order to get a complete evaluation of performance. The overall execution use from 1 to 8
-threads, and 4 to 1024 cores
+threads, and 4 to 2048 cores. Also we limit the total amount of threads to 2048 for cluster limitation, for a total of 34 runs, runned twice.
 
@@
 equal as the serial time.
-The reason for this approach is that a single core computation would require 1024 times the workload of the 1024 process version, but we calibrated the 1024 core version to take 3 minutes.
-
-\begin{tabular}{| c | c | c | c | c | c | c |}
+The reason for this approach is that a single core computation would require 1024 times the workload of the 1024 process version, but we calibrated the 1024 core version to take 3 minutes;
+so it should take around $3*1024/4=768min \approx 13h $, computing single core would have taken 4 times that.
+\begin{table}[H]
+\centering
+\tiny  % Use tiny font size for maximum space saving
+\setlength{\tabcolsep}{3pt}  % Reduce column separation
+\begin{tabular}{|c|*{10}{c}|}
   \hline
-  T/p & 4	 & 8	& 16	& 32	& 64 & 128  \\
+  T/P & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
   \hline
-  1 & 1	& 2,03 & 3,87 & 7,94 & 14,76 & 31,19  \\
-  2 & 1,04 & 2,05 & 3,97 & 8,50 & 16,43 & 27,90 \\	
-  4 & 1,01 &2,03	& 4,21 &	8,04 &	16,37 &	35,26	\\	
-  8 & 1,02  &2,02 &	4,02 &	8,62 &	18,32 &	32,26 \\
+  1 & -- & 14260.97 & 8471.77 & 3521.04 & 1849.25 & 1157.88 & 483.40 & 255.68 & 193.13 & 118.08 \\
+  2 & -- & 9659.36 & 5628.55 & 3259.06 & 2127.71 & 921.88 & 433.69 & 281.84 & 123.21 & -- \\
+  4 & 12785.30 & 7745.36 & 4652.30 & 2398.08 & 1465.56 & 687.89 & 453.28 & 252.71 & -- & -- \\
+  8 & 13781.07 & 11326.55 & 3800.54 & 2358.14 & 1308.72 & 722.74 & 597.09 & -- & -- & -- \\
   \hline
 \end{tabular}
-\newline\newline As we can see the speedup tends to 
-increase when the number of cores increase, however when
-we use more threads it happens that the
-speedup start to decrease at a certain point for some
-core configuration. As we can see from the plots it 
-seems there are threshold at some specific points, depending
-on the number of threads used; notice how the inclinations
-of the lines in the plot tends to be less and less evident.
+\end{table}
+
+As we can see the speedup increase almost linearly with the amount of processes used, in the efficency table it is shown that the efficency stays close to 1, until 512 process, then it deeps down.
+Probably it happens because the distance of the nodes is increasing (maybe throught a slower switch).
+Instead increasing thread parallelism is not as effective, and it seems to stagnate after 2 threads (but converging to 2)
+The reason could be that mpi pings cpu cores and process executions, so it could be limited to two (becuase Simultaneous Multithreading is usually set to 2).
 
@@
 the plot below it shows that there is a high variance in the
-datas, with the lines that tends to be more stretched when the
-number of threads gets higher
-\newline
-\newline
+\begin{table}[H]
+\centering
 \begin{tabular}{| c | c | c | c | c | c | c |}
@@
 \end{tabular}
+\caption{Efficiency results}
+\label{efficiency_results}
+\end{table}
 
@@
 \end{multicols}
+\section{Final execution and results}
+To sum up, the presented code is highly flexible and scalable, and it could be used for small optimization on low end systems as well as training other methods in an hpc architecture.
+The code is available to github, and it can be freely used in different applications.
 

### 2025-08-30 text-paper/paper.tex
tables

@@
 \centering
-\tiny  % Use tiny font size for maximum space saving
-\setlength{\tabcolsep}{3pt}  % Reduce column separation
+\tiny  
+\setlength{\tabcolsep}{3pt}
 \begin{tabular}{|c|*{10}{c}|}
@@
   \hline
-  1 & -- & 14260.97 & 8471.77 & 3521.04 & 1849.25 & 1157.88 & 483.40 & 255.68 & 193.13 & 118.08 \\
-  2 & -- & 9659.36 & 5628.55 & 3259.06 & 2127.71 & 921.88 & 433.69 & 281.84 & 123.21 & -- \\
-  4 & 12785.30 & 7745.36 & 4652.30 & 2398.08 & 1465.56 & 687.89 & 453.28 & 252.71 & -- & -- \\
-  8 & 13781.07 & 11326.55 & 3800.54 & 2358.14 & 1308.72 & 722.74 & 597.09 & -- & -- & -- \\
+  1 & 27879.62 & 14531.01 & 7700.45 & 3628.50 & 1873.90 & 1110.29 & 551.18 & 268.84 & 168.71 & 100.21 \\
+  2 & 33202.40 & 14243.06 & 5811.25 & 3274.74 & 1946.32 & 903.55 & 425.83 & 249.58 & 129.12 & -- \\
+  4 & 13136.02 & 8843.29 & 4623.91 & 2419.19 & 1460.78 & 724.75 & 461.13 & 260.10 & -- & -- \\
+  8 & 15110.34 & 11308.10 & 3850.66 & 2345.90 & 1281.55 & 737.66 & 541.53 & -- & -- & -- \\
   \hline
 \end{tabular}
+\caption{Time results}
+\label{time_results}
+\end{table}
+
+\begin{table}[H]
+\centering
+\tiny
+\setlength{\tabcolsep}{3pt}
+\begin{tabular}{|c|*{10}{c}|}
+  \hline
+  T/P & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
+  \hline
+  1 & 1 & 1.92 & 3.62 & 7.68 & 14.88 & 25.11 & 50.58 & 103.70 & 165.25 & 278.21 \\
+  2 & 0.84 & 1.96 & 4.80 & 8.51 & 14.32 & 30.86 & 65.47 & 111.71 & 215.93 & -- \\
+  4 & 2.12 & 3.15 & 6.03 & 11.52 & 19.09 & 38.47 & 60.46 & 107.19 & -- & -- \\
+  8 & 1.85 & 2.47 & 7.24 & 11.88 & 21.75 & 37.79 & 51.48 & -- & -- & -- \\
+  \hline
+\end{tabular}
+\caption{Speedup results}
+\label{speedup_results}
+\end{table}
+\begin{table}[H]
+\centering
+\tiny
+\setlength{\tabcolsep}{3pt}
+\begin{tabular}{|c|*{10}{c}|}
+  \hline
+  Efficiency & 4 & 8 & 16 & 32 & 64 & 128 & 256 & 512 & 1024 & 2048 \\
+  \hline
+  1 & 1 & 0.96 & 0.91 & 0.96 & 0.93 & 0.78 & 0.79 & 0.81 & 0.65 & 0.54 \\
+  2 & 0.42 & 0.49 & 0.60 & 0.53 & 0.45 & 0.48 & 0.51 & 0.44 & 0.42 & -- \\
+  4 & 0.53 & 0.39 & 0.38 & 0.36 & 0.30 & 0.30 & 0.24 & 0.21 & -- & -- \\
+  8 & 0.23 & 0.15 & 0.23 & 0.19 & 0.17 & 0.15 & 0.10 & -- & -- & -- \\
+  \hline
+\end{tabular}
+\caption{Efficiency results (normalized)}
+\label{efficiency_results_normalized}
 \end{table}
@@
 Probably it happens because the distance of the nodes is increasing (maybe throught a slower switch).
-Instead increasing thread parallelism is not as effective, and it seems to stagnate after 2 threads (but converging to 2)
-The reason could be that mpi pings cpu cores and process executions, so it could be limited to two (becuase Simultaneous Multithreading is usually set to 2).
+Instead increasing thread parallelism is not as effective, and it seems to give a $13\% $ improvment passing from 1 thread to 2 threads, $ 49\%$ to 4 threads, and $ 53\%$ to 8 threads.
+It seems that the thread level paralelization can be improved, however it seems that there is an hard limit, and we think it has to do with mpi core pinning.
+However the mpi scaling can replace openmp even on the same machine, providing better paralelism.
 

### 2025-08-30 text-paper/paper.tex
last revision

@@
   \textbf{
-    Dragonfly algorithm is an approach used to solve optimization problems through the emulation of dragonflies movements, but it lacks scalability.
-  In this paper the authors propose an approach based on parallelization using dedicated libraries for high performance multiprocess applications like MPI, and OpenMP API for multithread parallelization.
+    The dragonfly algorithm is an approach used to solve optimization problems through the emulation of dragonfly movements, but it lacks scalability.
+  In this paper, the authors propose an approach based on parallelization using dedicated libraries for high-performance multiprocess applications such as MPI, and the OpenMP API for multithread parallelization.
   A comparison of execution times is then conducted between presented implementation and the classic approach based on serial programming,
@@
 The Message Passing Interface (MPI) is a portable message-passing standard designed to function on parallel computing architectures.
-For our implementation it is usefull because it let us to express multiprocess comunication across multiple machines.
+For our implementation, it is useful because it allows us to express multiprocess communication across multiple machines.
 
 \subsection*{MPI on chunks}
-By far the biggest challenge in the parallelization with MPI was the comunication of the chunks partial computations (given that a chunk may or may not span multiple process).
-Mpi has an atomic function called $sendrecv$, that permits to send a message from a thread while receving from (possibly) another thread.
-This is super useful, but it could have caused deadlocks if every process has to comunicate with multiple chunks.
-So in order to use that we needed to ensure that each process is responsible of the accumulation of one chunk.
-We did this by ensuring that if a chunk starts in another process, and ends in this one, the process comunicate its contribute to the left process.
+By far the biggest challenge in the parallelization with MPI was the communication of the chunks' partial computations (given that a chunk may or may not span multiple processes).
+MPI has an atomic function called \textit{sendrecv}, that permits sending a message from a thread while receiving from (possibly) another thread.
+This is very useful, but it could have caused deadlocks if every process has to communicate with multiple chunks.
+So in order to use that, we needed to ensure that each process is responsible for the accumulation of one chunk.
+We did this by ensuring that if a chunk starts in another process and ends in this one, the process communicates its contribution to the left process.
 Then the accumulation is performed and the result is shared back to the current block.
@@
 \begin{itemize}
-  \item \textbf{Send left}: first, the process that have the end of a block, share their contribute to the left
-  \item \textbf{Accumulate and share} Then it is accumulated in the first process and broadcasted back to all partecipating processes
-  \item \textbf{Send right} Then the rightmost process in the block receives the accumulated value from the left.
+  \item \textbf{Send left}: First, the process that has the end of a block shares its contribution to the left.
+  \item \textbf{Accumulate and share}: Then it is accumulated in the first process and broadcast back to all participating processes.
+  \item \textbf{Send right}: Then the rightmost process in the block receives the accumulated value from the left.
 \end{itemize}
 
-In this way the comunication is efficient, and looking at the results it scales very wheel. More on that in section \ref{results}.
+In this way, the communication is efficient, and looking at the results, it scales very well. More on that in section \ref{results}.
 
@@
 
-OpenMP is a low level API that allows us to execute a sequence of instructions on different threads belonging
-to the same CPU core, improving in this way the performances of our implementation.
-Thanks to a compiler that supports this technology, in our case it is \textit{gcc},
+OpenMP is a low-level API that allows us to execute a sequence of instructions on different threads belonging
+to the same CPU core, thus improving the performance of our implementation.
+Thanks to a compiler that supports this technology (in our case, \textit{gcc}),
 we can include pragma directives in our code,
-which will wrap the set of instructions that we are interested to execute
+which will wrap the set of instructions that we are interested in executing
 in parallel. This feature can be used to 
-speed up the execution of some slow functions like the \textit{computation-accumulate}
+speed up the execution of some slow functions like the \textit{computation-accumulate}.
 
@@
 
-This function, after defining the starting and the ending points, compute
-the cumulated sums of dragonfiles positions and speeds 
-and saves the indexes of dragonfiles positions with the best fitness, using them
+This function, after defining the starting and ending points, computes
+the cumulated sums of dragonfly positions and speeds 
+and saves the indexes of dragonfly positions with the best fitness, using them
 to set the next locations of food and enemies.
@@
 
-The implementation of a parallelized approach in this case is obstacled by the fact that 
-threads should read and write on the same memeory areas like the \textit{cumulated\_pos},
-\textit{cumulated\_speed}, which could be 
+The implementation of a parallelized approach in this case is hindered by the fact that 
+threads should read and write to the same memory areas like the \textit{cumulated\_pos}
+and \textit{cumulated\_speed}, which could be 
 difficult to execute independently from each other due to race conditions. 
-In order to avoid this problem we split the current logical chunk into small ones, each
+In order to avoid this problem, we split the current logical chunk into smaller ones, each
 one of them keeps track of: 
@@
 \item the best local enemy and food fitnesses
-\item the indexes of the positions of dragonfiles with the best fitnesses
+\item the indexes of the positions of dragonflies with the best fitnesses
 \end{itemize}
-\noindent parallel executions takes care of a particular subchunk,
-defining also the initial and final positions by
-calculating the starting index as $thread\_ID*elems\_per\_thread$ and the end one as $start\_index+ratio$; 
-the final one however could be responsible for more dragonfiles 
+\noindent Parallel executions take care of a particular subchunk,
+also defining the initial and final positions by
+calculating the starting index as $thread\_ID \times elems\_per\_thread$ and the end one as $start\_index + ratio$; 
+the final thread, however, could be responsible for more dragonflies 
 than others since the number of threads does not always
-divide the number of elements perfectly. The single threads then proceed to 
+divide the number of elements perfectly. The individual threads then proceed to 
 calculate the cumulated sums and the fitnesses of the current dragonfly, 
-saving its position index if it maximize the food fitness or minimize the 
+saving its position index if it maximizes the food fitness or minimizes the 
 enemy fitness. 
@@
  \end{lstlisting}
- After the join of all threads, the program performs a reduce of subchunks, calculating
-the optimal . This methodology can be seen as a unique critical section
-executed only once after all the parallel processings have ended; then as final step the position and fitnesses
-of food and enemies is updated, without the need to define critical sections which would force the processor 
+ After the join of all threads, the program performs a reduction of subchunks, calculating
+the optimal values. This methodology can be seen as a single critical section
+executed only once after all the parallel processing has ended; then as a final step, the positions and fitnesses
+of food and enemies are updated, without the need to define critical sections which would force the processor 
 to update the position buffers sequentially.
@@
 \section{Final execution and results}
-Now that we have described all the procedure it is time to illustrate how 
+Now that we have described all the procedures, it is time to illustrate how 
 the algorithm has been executed. This job has been performed by the
-High Performance Computing cluster at University of Trento, which 
-use a particular technology called PBS to handle multiple tasks at the same
-time: this feature comes useful in our case since, as we will see later, we 
-execute multiple istances of the same problem
+High Performance Computing cluster at the University of Trento, which 
+uses a particular technology called PBS to handle multiple tasks at the same
+time: this feature is useful in our case since, as we will see later, we 
+execute multiple instances of the same problem.
 \subsection*{Structure of the execution}
 
-As anticipated we will execute the same problem using multiple instances with different amount of
-resources in order to get a complete evaluation of performance. The overall execution use from 1 to 8
-threads, and 4 to 2048 cores. Also we limit the total amount of threads to 2048 for cluster limitation, for a total of 34 runs, runned twice.
+As anticipated, we will execute the same problem using multiple instances with different amounts of
+resources in order to get a complete evaluation of performance. The overall execution uses from 1 to 8
+threads and 4 to 2048 cores. Also, we limit the total number of threads to 2048 for cluster limitations, for a total of 34 runs, run twice.
 
@@
 
-After the end of all tasks it is possible to check how much gain in performance we have obtained.
-To do this it is necessary to introduce some theoretical concepts that will give us the tools that
-will give us a good metric system to evaluate our results.Speedup is defined as the ratio between the the execution times of the serial and parallel
-implementations, so the formula will be equal to: $$Speedup = \frac{T_{serial}}{T_{parallel}}$$ 
-from this definition it is obvious that the more the speedup is the more the parallel implementation
-performs better. There is a particular case where the time execution of the denominator can be defined
+After the end of all tasks, it is possible to check how much gain in performance we have obtained.
+To do this, it is necessary to introduce some theoretical concepts that will provide us with the tools 
+to give us a good metric system to evaluate our results. Speedup is defined as the ratio between the execution times of the serial and parallel
+implementations, so the formula will be: $$Speedup = \frac{T_{serial}}{T_{parallel}}$$ 
+From this definition, it is obvious that the greater the speedup, the better the parallel implementation
+performs. There is a particular case where the execution time of the denominator can be defined
 as $T_{parallel}=T_{serial}/p$, where \textit{p} is the number of cores: this is called \textit{linear speedup},
-and in this situation it perfectly match with the classic speedup coefficient
+and in this situation it perfectly matches the classic speedup coefficient.
 Efficiency is another important parameter, and it is defined by the ratio of the speedup and the 
 number of cores, giving the following formula: $$E = \frac{Speedup}{p}$$
-also in this case the efficiency decrease due to the increase of the number of cores. In case
-of linear speedup we get that the efficiency is equal to the inverse of $p$
+Also in this case, the efficiency decreases due to the increase in the number of cores. In the case
+of linear speedup, we get that the efficiency is equal to the inverse of $p$.
 \subsection*{results}
 \label{results}
-Now it is possible to show results obtained from the cluster,
-using the configuration with 4 cores and 1 thread will be considered
-equal as the serial time.
+Now it is possible to show the results obtained from the cluster.
+Using the configuration with 4 cores and 1 thread will be considered
+equal to the serial time.
 The reason for this approach is that a single core computation would require 1024 times the workload of the 1024 process version, but we calibrated the 1024 core version to take 3 minutes;
-so it should take around $3*1024/4=768min \approx 13h $, computing single core would have taken 4 times that.
+so it should take around $3 \times 1024/4 = 768 \textrm{ min} \approx 13\textrm{h}$. Computing with a single core would have taken 4 times that.
 \begin{table}[H]
@@
 
-As we can see the speedup increase almost linearly with the amount of processes used, in the efficency table it is shown that the efficency stays close to 1, until 512 process, then it deeps down.
-Probably it happens because the distance of the nodes is increasing (maybe throught a slower switch).
-Instead increasing thread parallelism is not as effective, and it seems to give a $13\% $ improvment passing from 1 thread to 2 threads, $ 49\%$ to 4 threads, and $ 53\%$ to 8 threads.
-It seems that the thread level paralelization can be improved, however it seems that there is an hard limit, and we think it has to do with mpi core pinning.
-However the mpi scaling can replace openmp even on the same machine, providing better paralelism.
+As we can see, the speedup increases almost linearly with the number of processes used.
+In the efficiency table, it is shown that the efficiency stays close to 1 until 512 processes, then it drops down.
+This probably happens because the distance between the nodes is increasing (maybe through a slower switch).
+Instead, increasing thread parallelism is not as effective, and it seems to give a $13\%$ improvement when passing from 1 thread to 2 threads, $49\%$ to 4 threads, and $53\%$ to 8 threads.
+It seems that the thread-level parallelization can be improved; however, it seems that there is a hard limit, and we think it has to do with MPI core pinning.
+However, the MPI scaling can replace OpenMP even on the same machine, providing better parallelism.
 
@@
 \noindent 
-Efficiency tends to present an unpredictable behaviour,
-in fact the coefficients grows for some particular setups,
-for others instead numbers decrease. If we take a look on
-the plot below it shows that there is a high variance in the
+Efficiency tends to present unpredictable behavior.
+In fact, the coefficients grow for some particular setups,
+while for others, the numbers decrease. If we take a look at
+the plot below, it shows that there is high variance in the results.
 \begin{table}[H]
@@
 \section{Final execution and results}
-To sum up, the presented code is highly flexible and scalable, and it could be used for small optimization on low end systems as well as training other methods in an hpc architecture.
-The code is available to github, and it can be freely used in different applications.
+To sum up, the presented code is highly flexible and scalable, and it could be used for small optimizations on low-end systems as well as training other methods in an HPC architecture.
+The code is available on GitHub, and it can be freely used in different applications.
 

### 2025-08-30 text-paper/paper.tex
put matricola IDs

@@
 \title{\textbf{Parallel dragonfly algorithm} %
-  \\[2ex] \large High Performance Computing for Data Science course 2024/2025}
+  \\[2ex] \large High Performance Computing for Data Science course 2024/2025
+}
 
-\author[1]{Mattia Santaniello}
-\author[2]{Alessio Zeni}
-\affil{University of Trento}
+\author{
+  Mattia Santaniello \quad  mat. 258777 \\ Alessio Zeni  \quad
+  mat. 256130  
+}
 
@@
 To sum up, the presented code is highly flexible and scalable, and it could be used for small optimizations on low-end systems as well as training other methods in an HPC architecture.
-The code is available on GitHub, and it can be freely used in different applications.
+The code is available on GitHub, and it can be freely used in different applications.\cite{GitHub}
 

### 2025-08-30 text-paper/paper.tex
updated graphics
conclusions

@@
 \begin{figure}[H]
-  \includegraphics[scale=0.3]{img/speedup.jpg}
+  \includegraphics[scale=0.5]{img/speedup.jpg}
   \centering
-  \caption{speedup graphic}
+  \caption{Speedup graphic}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[scale=0.3]{img/efficiency.jpg}
+  \includegraphics[scale=0.5]{img/efficency.jpg}
   \centering
-  \caption{efficiency graphic}
+  \caption{Efficency graphic}
 \end{figure}
@@
 \end{multicols}
-\section{Final execution and results}
-To sum up, the presented code is highly flexible and scalable, and it could be used for small optimizations on low-end systems as well as training other methods in an HPC architecture.
-The code is available on GitHub, and it can be freely used in different applications.\cite{GitHub}
+\section{Conclusion and final considerations}
+The presented code is highly flexible and scalable, and it could be used for small optimizations on low-end systems as well as 
+training other methods in an HPC architecture. There are some considerations however that need to be cited and discussed:
+
+\begin{itemize}
+
+\item Further analysis should be conducted in order to determine if there are better approaches that could surpass the proposed one.
+As shown in the previous section OpenMP contribute does not seem fundamental for the increase of performance, suggesting a possible bottleneck
+that should be optimized, peraphs considernig other technologies and parallelizing other functions than \textit{computation-accumulate}.
+
+\item Another relevant aspect that need to be investigated is the test of fitness functions different from Rastrigin, looking at how they behave
+in different situations with a HPC approach, considering also some real application in biology and engineering;
+however it is pretty reasonable to assume that results should not differ too much from the ones illustrated in this paper since
+Rastrigin is a hard function to optimize.
+\end{itemize} The code is available on GitHub, and it can be freely 
+used in different applications.\cite{GitHub}
 
